---
title: "behavioral analysis"
author: "Christopher"
date: "16 6 2025"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 5
    theme: flatly
    df_print: paged
---



```{r load_package, message = FALSE, warning = FALSE}
library(lme4)
library(readxl)
library(tidyverse)
# library(data.table)
# library(afex)
# library(broom)
# library(apa)
library(kableExtra)
# library(lmerTest)
# library(pbkrtest)
# library(effects)
# library(visreg)
# library(sjPlot)
library(broom.mixed)
# library(pander)
# library(mediation)
# library(multcomp)
# library(multcompView)
# library(magrittr)
library(multipanelfigure)
library(ggbeeswarm)
# library(lsmeans)
library(BayesFactor)
library(ggpubr)
# library(gpairs)
# library(DescTools)
library(cowplot)
# library(ggpol)
# library(psych)
library(ggdist)
# library(dplyr)
# 
library(mgcv)
library(tidyverse)       # Tidy and flexible data manipulation
# library(marginaleffects) # Compute conditional and marginal effects
# library(patchwork)       # Combining ggplot objects
# library(gratia)
# 
library(ggeffects)
# 
library(report)
library(emmeans)


source('C:/Dropboxdata/Dropbox/work/R-statistics/general_functions/RainCloudPlots/tutorial_R/R_rainclouds.R')





# broad cluster
# CSD based normalized to pre-cue baseline in %
# 
DATAPath1 = "data_in/behavior_events.csv"
DATAPath2 = "data_in/behavior_FAs.csv"





options(scipen=1, digits=5)
```
<style type="text/css">
.main-container {
  max-width: 1800px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>


## Prepare data  
<br>  

1.  Read in data: 
    + `r DATAPath1`
    + `r DATAPath1`
    
<br>  

```{r load_data,results = "hide", fig.show = "hide", warning = FALSE}
# read in data
DataIn1 <- read_csv(DATAPath1)
DataIn2 <- read_csv(DATAPath2)
head(DataIn1)
str(DataIn1)
head(DataIn2)
str(DataIn2)

```





1.  Prepare new dataset(s)...
<br>  

```{r prepare_data1, results = "hide", fig.show = "hide", warning = FALSE}
DATA_Trials <- DataIn1%>%
  filter(eventtype=="target")


```

## How to cite the packages?

<br> 
```{r cite data, results = "hide", fig.show = "hide", warning = FALSE}
session <- sessionInfo()
r <- report(session)
r
summary(r)
as.data.frame(r)
summary(as.data.frame(r))

```

## Illustrate behavioral data

<br>  

```{r plot_data_avg, results = "hide",  fig.height=6, fig.width=7, warning = FALSE}

dat2plot1 <- DATA_Trials %>%
  group_by(participant, eventtype2)%>%
  dplyr::summarise(
    RT_Mean=mean(RT,na.rm=TRUE),
    RT_STD=sd(RT,na.rm=TRUE),
    Hitrate=(sum(response=='hit')/n())*100,
    Hitrate_adj1 = ifelse((sum(response=='hit')/n())>0,
                         min(c(sum(response=='hit')/n(), 1-(.5/n()))),
                         .5/n()), # according to Stanislaw H, Todorov N, Behav Res Meth (1999) 31, 137-149, "1/2N rule" or
    Hitrate_adj2 = (sum(response=='hit')+0.5)/(n()+1))%>% # loglinear approach (Hautus, 1995)
    ungroup()
dat2plot1a = DataIn1%>%
  filter(eventtype=="distractor") %>%
  group_by(participant, eventtype2)%>%
  dplyr::summarise(
    FArate=(sum(response=='FA_proper')/n())*100,
    FArate_adj1 = ifelse((sum(response=='FA_proper')/n())>0,
                         min(c(sum(response=='FA_proper')/n(), 1-(.5/n()))),
                         .5/n()), # according to Stanislaw H, Todorov N, Behav Res Meth (1999) 31, 137-149, "1/2N rule" or
    FArate_adj2 = (sum(response=='FA_proper')+0.5)/(n()+1)) %>% # loglinear approach (Hautus, 1995))
  ungroup()
dat2plot <- merge(dat2plot1,dat2plot1a,by = "participant") %>% # adjust for HR and FR of 0 or 1
  mutate(
    d_prime = qnorm(Hitrate_adj1)-qnorm(FArate_adj1),
    c = -0.5*(qnorm(Hitrate_adj1)+qnorm(FArate_adj1))
  ) %>%
  dplyr::rename(eventtype2=eventtype2.x)




dat2plot2 <- DataIn2 %>%
  mutate(`FA_all`= FA + `FA_proper`)%>%
  # dplyr::select(-FA)%>%
  dplyr::select(-FA_all)%>%
  pivot_longer(c(-subject ),names_to = "FA_type", values_to = "number")%>%
  filter(FA_type=='FA_proper' | FA_type=='FA')

theme_set(theme_bw())
plot1 <- ggplot(dat2plot, aes(x = 1, y = RT_Mean)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  ylab(expression(paste("reaction time in ms")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

plot2 <- ggplot(dat2plot, aes(x = 1, y = Hitrate)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange"))
  ylab(expression(paste("hit rate in %")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

plot3 <- ggplot(dat2plot2, aes(x = FA_type, y = number, color = FA_type)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=3.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  scale_fill_manual("FA type", values=c("grey40", "black"),labels = c("to other","to distractor")) +
  scale_color_manual("FA type", values=c("grey40", "black"),labels = c("to other","to distractor")) +
  ylab(expression(paste("number of False Alarms")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

plot4 <- ggplot(dat2plot, aes(x = 1, y = d_prime)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange"))
  ylab(expression(paste("dprime")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())


# cowplot::plot_grid(plot1, plot2, plot3, ncol = 3,labels = "AUTO")
plotttitle <- ggdraw() + draw_label("behavioral measures", fontface='bold')
plotplots <- ggarrange(plot1, plot2, plot3, plot4, ncol = 4, align = "h", widths=c(2,2,5,2))
ggarrange(plotttitle, plotplots, nrow = 2, heights=c(0.2, 1))




## for publication
theme_set(theme_bw())
plot1 <- ggplot(dat2plot, aes(x = eventtype2, y = RT_Mean, color =eventtype2,)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  ylab(expression(paste("reaction time in ms")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_y_continuous(limits=c(400, 775))

plot1b <- ggplot(dat2plot, aes(x = 1, y = RT_Mean, color =eventtype2, fill = eventtype2)) +
  geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.0, trim = FALSE, alpha = 0.8)+
  # scale_fill_manual(values=c("grey40"))+
  scale_fill_brewer(palette = "Dark2")+
  # scale_color_manual(values=c("grey40"))+
  scale_color_brewer(palette = "Dark2")+
  theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
         axis.ticks.y=element_blank())+
  theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
         axis.ticks.x=element_blank())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_y_continuous(limits=c(400, 775))

plot2 <- ggplot(dat2plot, aes(x = eventtype2, y = Hitrate, color =eventtype2)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
    # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange"))
  ylab(expression(paste("hit rate in %")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_y_continuous(limits=c(30, 100))

plot2b <- ggplot(dat2plot, aes(x = 1, y = Hitrate,  color =eventtype2, fill = eventtype2)) +
  geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.0, trim = FALSE, alpha = 0.8)+
  # scale_fill_manual(values=c("grey40"))+
  scale_fill_brewer(palette = "Dark2")+
  # scale_color_manual(values=c("grey40"))+
  scale_color_brewer(palette = "Dark2")+
  theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
         axis.ticks.y=element_blank())+
  theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
         axis.ticks.x=element_blank())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_y_continuous(limits=c(30, 100))

  

# cowplot::plot_grid(plot1, plot2, plot3, ncol = 3,labels = "AUTO")
plotttitle <- ggdraw() + draw_label("behavioral measures", fontface='bold')
plotplots <- ggarrange(plot1, plot1b,plot2, plot2b, ncol = 4,common.legend = TRUE,align = "h", widths=c(3,2,3,2))

savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", width = 5, height = 6, unit = "in")
saveplot <- ggarrange(plotttitle, plotplots, nrow = 2, heights=c(0.2, 1))

savefig %<>%
  fill_panel(saveplot, column = 1, row = 1)
savefig
# save_multi_panel_figure(savefig, "hits_reactiontimes.eps")


## for publication
theme_set(theme_bw())
plot1 <- ggplot(dat2plot, aes(x = eventtype2, y = RT_Mean, color =eventtype2)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  ylab(expression(paste("reaction time in ms")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_y_continuous(limits=c(400, 775))

plot1b <- ggplot(dat2plot, aes(x = 1, y = RT_Mean, color =eventtype2, fill = eventtype2)) +
  geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.0, trim = FALSE, alpha = 0.8)+
  # scale_fill_manual(values=c("grey40"))+
  scale_fill_brewer(palette = "Dark2")+
  # scale_color_manual(values=c("grey40"))+
  scale_color_brewer(palette = "Dark2")+
  theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
         axis.ticks.y=element_blank())+
  theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
         axis.ticks.x=element_blank())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_y_continuous(limits=c(400, 775))

plot2 <- ggplot(dat2plot, aes(x = eventtype2, y = d_prime, color =eventtype2)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange"))
  ylab(expression(paste("dprime")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_y_continuous(limits=c(0.0, 3.8))

plot2b <- ggplot(dat2plot, aes(x = 1, y = d_prime, color =eventtype2, fill = eventtype2)) +
  geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.0, trim = FALSE, alpha = 0.8)+
  # scale_fill_manual(values=c("grey40"))+
  scale_fill_brewer(palette = "Dark2")+
  # scale_color_manual(values=c("grey40"))+
  scale_color_brewer(palette = "Dark2")+
  theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
         axis.ticks.y=element_blank())+
  theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
         axis.ticks.x=element_blank())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_y_continuous(limits=c(0.0, 3.8))

  

# cowplot::plot_grid(plot1, plot2, plot3, ncol = 3,labels = "AUTO")
plotttitle <- ggdraw() + draw_label("behavioral measures", fontface='bold')
plotplots <- ggarrange(plot1, plot1b,plot2, plot2b, ncol = 4,common.legend = TRUE,align = "h", widths=c(3,2,3,2))

savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", width = 5, height = 6, unit = "in")
saveplot <- ggarrange(plotttitle, plotplots, nrow = 2, heights=c(0.2, 1))

savefig %<>%
  fill_panel(saveplot, column = 1, row = 1)
savefig
# save_multi_panel_figure(savefig, "dprime_reactiontimes.eps")
  

```


## Display behavioral data numerically

<br>  

```{r disp_data_avg, warning = FALSE}
dat2plot1 <- DATA_Trials %>%
  group_by(eventtype2, participant )%>%
  dplyr::summarise(
    RT_Mean=mean(RT,na.rm=TRUE),
    RT_STD=sd(RT,na.rm=TRUE),
    Hitrate=(sum(response=='hit')/n())*100,
    Hitrate_adj1 = ifelse((sum(response=='hit')/n())>0,
                         min(c(sum(response=='hit')/n(), 1-(.5/n()))),
                         .5/n()), # according to Stanislaw H, Todorov N, Behav Res Meth (1999) 31, 137-149, "1/2N rule" or
    Hitrate_adj2 = (sum(response=='hit')+0.5)/(n()+1)) # loglinear approach (Hautus, 1995)

dat2plot1a = DataIn1%>%
  filter(eventtype=="distractor") %>%
  group_by(eventtype2, participant)%>%
  dplyr::summarise(
    FArate=(sum(response=='FA_proper')/n())*100,
    FArate_adj1 = ifelse((sum(response=='FA_proper')/n())>0,
                         min(c(sum(response=='FA_proper')/n(), 1-(.5/n()))),
                         .5/n()), # according to Stanislaw H, Todorov N, Behav Res Meth (1999) 31, 137-149, "1/2N rule" or
    FArate_adj2 = (sum(response=='FA_proper')+0.5)/(n()+1)) # loglinear approach (Hautus, 1995))

dat2plot <- merge(dat2plot1,dat2plot1a,by = "participant") %>% # adjust for HR and FR of 0 or 1
  mutate(
    d_prime = qnorm(Hitrate_adj1)-qnorm(FArate_adj1),
    c = -0.5*(qnorm(Hitrate_adj1)+qnorm(FArate_adj1))
  ) %>%
  dplyr::rename(eventtype2=eventtype2.x)

dat2plot  %>%
  group_by(eventtype2)%>%
  dplyr::summarise(N = n(), mean_RT = mean(RT_Mean), std_RT = sd(RT_Mean), mean_Hitrate = mean(Hitrate), std_Hitrate = sd(Hitrate),
            mean_FArate = mean(FArate), std_FArate = sd(FArate), mean_dprime = mean(d_prime), std_dprime = sd(d_prime)) %>%
  kable(escape = F, digits = c(3,3,3,3,3,3,3,3,3,3), caption = c("descriptives of behavioral measures")) %>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8))

dat2plot  %>%
  group_by(participant)%>%
  dplyr::summarise(N = n(), mean_RT = mean(RT_Mean), std_RT = sd(RT_Mean), mean_Hitrate = mean(Hitrate), std_Hitrate = sd(Hitrate),
            mean_FArate = mean(FArate), std_FArate = sd(FArate), mean_dprime = mean(d_prime), std_dprime = sd(d_prime)) %>%
  ungroup()%>%
  dplyr::summarise(N = n(), M_RT = mean(mean_RT), SD_RT = sd(mean_RT), M_Hitrate = mean(mean_Hitrate), SD_Hitrate = sd(mean_Hitrate),
            M_FArate = mean(mean_FArate), SD_FArate = sd(mean_FArate), M_dprime = mean(mean_dprime), SD_dprime = sd(mean_dprime)) %>%
  kable(escape = F, digits = c(3,3,3,3,3,3,3,3,3,3), caption = c("descriptives of behavioral measures")) %>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8))

```



## Statistical analysis

### ANOVA models
### linear mixed models

#### test RT, Hitrate and dprime between primed and non primed targets

**RT** ~ **priming**

<br> 

```{r lme_RT_lum, fig.show = "hide", warning = FALSE, fig.height=4, fig.width=10}
dat2plot2 <- DataIn2 %>%
  mutate(`FA_all`= FA + `FA_proper`)%>%
  # dplyr::select(-FA)%>%
  dplyr::select(-FA_all)%>%
  pivot_longer(c(-subject ),names_to = "FA_type", values_to = "number")%>%
  filter(FA_type=='FA_proper' | FA_type=='FA')


dat2plot1 <- DATA_Trials %>%
  group_by(participant,eventtype2)%>%
  dplyr::summarise(RT_Mean=mean(RT,na.rm=TRUE),RT_STD=sd(RT,na.rm=TRUE),Hitrate=(sum(response=='hit')/n())*100)

dat2plot2 <- DataIn2 %>%
  mutate(`FA_all`= FA + `FA_proper`)%>%
  # dplyr::select(-FA)%>%
  dplyr::select(-FA_all)%>%
  dplyr::rename(participant=subject)

dat2plot_all <- merge(dat2plot1,dat2plot2,by = "participant") %>% # adjust for HR and FR of 0 or 1
  mutate(
    d_prime = qnorm(Hitrate)-qnorm(FA_proper_rate),
    c = -0.5*(qnorm(Hitrate)+qnorm(FA_proper_rate))
  )



# # statisticical analysis ANOVA
# StatsOut <- dat2plot_all %>%
#   aov_ez(id="participant", dv = "RT_Mean", data = ., between = c("colorlum"), include_aov = afex_options("include_aov"))
# 
# StatsOut%>%
#   .$anova %>%
#   mutate(factor = rownames(.))%>%
#   mutate(
#     `Pr(>F)` = cell_spec(ifelse(round(`Pr(>F)`,4)<.001,"< .001",round(`Pr(>F)`,4)),
#                         color = ifelse(is.nan(`Pr(>F)`),"blue", ifelse(`Pr(>F)` < .05, "green", "red")),
#                         bold = ifelse(is.nan(`Pr(>F)`),F, ifelse(`Pr(>F)` < .05, T,F)),
#                         align = "center")
#   ) %>%
#   dplyr::select(`num Df`, `den Df`, MSE, `F`, ges, `Pr(>F)`) %>%
#   kable(escape = F, digits = c(3,3,3,3,3,3,3,3,3), caption = c("ANOVA | dv = RT in ms")) %>%
#   kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 6))


# RT
dat2plot_all %>%
  nest() %>%
  mutate(
    # stats = map(data, ~ broom::tidy(t.test(RT_Mean ~ eventtype2, data = ., paired = TRUE))),
    stats = map(data, ~ rstatix::t_test(RT_Mean ~ eventtype2, data = ., paired = TRUE, detailed = TRUE)),
    t_test_bf = map(data, ~ extractBF(ttestBF(formula = RT_Mean ~ eventtype2, data = ., 
                                              iterations = num.iter, rscale = sqrt(2)/2))),
    meanval = map(data, ~ .x %>%
                    group_by(eventtype2) %>%
                    summarise(
                      mean_RT_Mean = mean(RT_Mean, na.rm = TRUE),
                      std_RT_Mean = sd(RT_Mean, na.rm = TRUE)
                    ) %>%
                    pivot_wider(names_from = eventtype2, 
                                values_from = c(mean_RT_Mean, std_RT_Mean), 
                                names_sep = "_")
    ),
    Dz = map(data, ~ rstatix::cohens_d(RT_Mean ~ eventtype2, data = ., paired = TRUE))
  ) %>%
  # unnest(cols = c(stats, cohens_d, t_test_bf, meanval))%>%
  unnest(cols = c(Dz))%>%
  dplyr::select(stats,t_test_bf, meanval, effsize) %>%
  unnest(cols = c(stats, t_test_bf, meanval)) %>%
  mutate(BF10 = bf, BF01 = 1/bf, 
         M_RT_noprim = mean_RT_Mean_target_nonprimed,
         SD_RT_noprim = std_RT_Mean_target_nonprimed,
         M_RT_prim = mean_RT_Mean_target_primed,
         SD_RT_prim = std_RT_Mean_target_primed,
         diff=estimate,
         Dz = effsize, t.value=statistic)%>%
  dplyr::select(M_RT_noprim, SD_RT_noprim, M_RT_prim, SD_RT_prim, diff, t.value, df, p, Dz, BF10, BF01)%>%
  # mutate(p = p.adjust(p, method="holm")) %>% # none, holm
  # mutate(p = p.adjust(p, method="none")) %>% # none, holm
  ungroup()%>%
  # mutate(
  #   p = cell_spec(p, color = ifelse(is.nan(p),"blue", ifelse(p < .05, "green", "red")))
  # ) %>%
  mutate(
    p = cell_spec(ifelse(round(p,4)<.001,"< .001",round(p,4)),
                  color = ifelse(is.nan(p),"blue", ifelse(p < .05, "green", "red")),
                  bold = ifelse(is.nan(p),F, ifelse(p < .05, T,F)),
                  align = "center")
  ) %>%
  kbl(escape = F, digits = 3, 
        caption = c("RT | modulation by target type (primed vs nonprimed) | t-tests of difference")) %>%
  kable_styling("striped", full_width = T,bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8))
```

**Hitrate** ~ **priming**

<br> 

```{r lme_HR_lum, fig.show = "hide", warning = FALSE, fig.height=4, fig.width=10}
# Hitrate
dat2plot_all %>%
  nest() %>%
  mutate(
    # stats = map(data, ~ broom::tidy(t.test(Hitrate ~ eventtype2, data = ., paired = FALSE))),
    stats = map(data, ~ rstatix::t_test(Hitrate ~ eventtype2, data = ., paired = TRUE, detailed = TRUE)),
    t_test_bf = map(data, ~ extractBF(ttestBF(formula = Hitrate ~ eventtype2, data = ., iterations = num.iter, rscale = sqrt(2)/2))),
    meanval = map(data, ~ .x %>%
                    group_by(eventtype2) %>%
                    summarise(
                      mean_HR_Mean = mean(Hitrate, na.rm = TRUE),
                      std_HR_Mean = sd(Hitrate, na.rm = TRUE)
                    ) %>%
                    pivot_wider(names_from = eventtype2, 
                                values_from = c(mean_HR_Mean, std_HR_Mean), 
                                names_sep = "_")
    ),
    cohens_d = map(data, ~ rstatix::cohens_d(Hitrate ~ eventtype2, data = ., paired = FALSE))
  ) %>%
  # unnest(cols = c(stats, cohens_d, t_test_bf, meanval))%>%
  unnest(cols = c(cohens_d))%>%
  dplyr::select(stats,t_test_bf, meanval, effsize) %>%
  unnest(cols = c(stats, t_test_bf, meanval)) %>%
  mutate(BF10 = bf, BF01 = 1/bf, 
         M_HR_noprim = mean_HR_Mean_target_nonprimed,
         SD_HR_noprim = std_HR_Mean_target_nonprimed,
         M_HR_prim = mean_HR_Mean_target_primed,
         SD_HR_prim = std_HR_Mean_target_primed,
         diff=estimate,
         Dz = effsize, t.value=statistic)%>%
  dplyr::select(M_HR_noprim, SD_HR_noprim, M_HR_prim, SD_HR_prim, diff, t.value, df, p, Dz, BF10, BF01)%>%
  # mutate(p = p.adjust(p, method="holm")) %>% # none, holm
  # mutate(p = p.adjust(p, method="none")) %>% # none, holm
  ungroup()%>%
  mutate(
    p = cell_spec(ifelse(round(p,4)<.001,"< .001",round(p,4)),
                        color = ifelse(is.nan(p),"blue", ifelse(p < .05, "green", "red")),
                        bold = ifelse(is.nan(p),F, ifelse(p < .05, T,F)),
                        align = "center")
  ) %>%
  kbl(escape = F, digits = 3, 
        caption = c("Hitrate | modulation by target type (primed vs nonprimed) | t-tests of difference")) %>%
  kable_styling("striped", full_width = T,bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8))
```
<br> 

## Time-resolved Analysis

### Reaction Times

<br> 

#### Statistical test for time-resolved analysis of probability of hits/hit rates
*Do probabilities of Hits for primed and non-primed events differ (differently) across time?*

full model: response ~ eventtype2 * eventonsettimes + (1 | participant)
reduced model: response ~ eventtype2 + eventonsettimes + (1 | participant)

Test rationale: do these models differ significantly and does the interaction term thus allow for a better explanation of the data.

<br> 
```{r stat_hitxpresentationtime_distr, warning = FALSE, fig.height=4, fig.width=6}
# Full model with interaction

# Model with smoothers by condition
# with time x cue interaction
model_gam_interaction <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         eventtype2 = as.factor(eventtype2))%>%
  gam(hit ~ eventtype2 + s(eventonsettimes, by = eventtype2, bs = "cs") + participant,
                 data = ., family = binomial)
  #
# coef(model_gam_interaction)
# summary(model_gam_interaction)

# no interaction only main effects
model_gam_TwoMain <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         eventtype2 = as.factor(eventtype2))%>%
  gam(hit ~ eventtype2 + s(eventonsettimes, bs = "cs") + participant, 
                 data = ., family = binomial)

# no effect of time only CUE
model_gam_OneMain <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         eventtype2 = as.factor(eventtype2))%>%
  gam(hit ~ eventtype2  + participant, 
                 data = ., family = binomial)

# no effect of time
model_gam_null <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         eventtype2 = as.factor(eventtype2))%>%
  gam(hit ~  participant, 
                 data = ., family = binomial)

plot(model_gam_interaction, pages = 1)
# anova(model_gam_interaction)


# based on 'marginal effects'
# plot_predictions(model_gam_interaction, condition = c('eventonsettimes', 'eventtype2'),
#                  type = 'response') +
#       labs(y = "Predicted probability of a hit",
#            title = "Average smooth effect of eventonsettimes | GAM",
#            subtitle = "Per priming condition")


anova(model_gam_null,model_gam_OneMain,model_gam_TwoMain,model_gam_interaction) %>%
  tidy()%>%
  kbl(escape = F, digits=3,
      caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:"))%>%
  # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>",
  #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = 
                                                                   "hold_position"))


AIC(model_gam_null, model_gam_OneMain, model_gam_TwoMain, model_gam_interaction) %>%
  # tidy()%>%
  kbl(escape = F, digits=3,
      caption = paste0("AIC | comparing full and reduced model | GAM:"))%>%
  # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>",
  #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options =
                                                                   "hold_position"))
# based on 'gratia'
# # Plot difference between smooths (e.g., eventtype2 levels "A" and "B")
# diff_df <- difference_smooths(model_gam_interaction,
#                                smooth = "s(eventonsettimes)",
#                                comparison = c("eventtype2A", "eventtype2B"),
#                                n = 10)  # 200 points across the range

# diff_df <- diff_df %>%
#   mutate(
#     .diff = plogis(.diff),
#     .lower_c = plogis(.lower_ci),
#     .upper_ci = plogis(.upper_ci)
#   )

# ggplot(diff_df, aes(x = eventonsettimes, y = .diff)) +
#   geom_line() +
#   geom_ribbon(aes(ymin = .lower_ci, ymax = .upper_ci), alpha = 0.3) +
#   geom_hline(yintercept = 0, linetype = "dashed") +
#   labs(x = "Event onset time", y = "Difference in log-odds of Hitrate (nonprimed - primed)")


```

<br> 

##### follow up: directly test time windows | Hitrates
*Do probabilities of Hits for primed and non-primed events differ (differently) across time?*

full model: response ~ eventtype2 * eventonsettimes + (1 | participant)

Test rationale: do these models differ significantly and does the interaction term thus allow for a better explanation of the data.

Extract predicted probability of hits for early and late window and contrast them

<br> 

```{r stat_hitxpresentationtime_distr_contrasts, warning = FALSE, fig.height=3.5, fig.width=4.9}
# Full model with interaction
model_gam_interaction <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         eventtype2 = as.factor(eventtype2))%>%
  gam(hit ~ eventtype2 + s(eventonsettimes, by = eventtype2, bs = "cs") + participant,
                 data = ., family = binomial)

# alternative approach
# Define the time windows
all_timepoints = sort(unique(DATA_Trials$eventonsettimes))
early_window <- all_timepoints[
  (all_timepoints >= min(unique(DATA_Trials$eventonsettimes)) + 0) & (all_timepoints <= min(unique(DATA_Trials$eventonsettimes)) + 0.5)
]
late_window <-all_timepoints[
  (all_timepoints >= max(unique(DATA_Trials$eventonsettimes)) -.5) & (all_timepoints <= max(unique(DATA_Trials$eventonsettimes)) + 0)
]

emm_predictions_early <- emmeans(
  model_gam_interaction,
  ~ eventtype2 ,
  at = list(eventonsettimes = early_window)
)

emm_predictions_late <- emmeans(
  model_gam_interaction,
  ~ eventtype2 ,
  at = list(eventonsettimes = late_window)
)

# 1. Add the identifier column to the grid of each object
emm_predictions_early@grid$eventonsettimes <- "early"
emm_predictions_late@grid$eventonsettimes  <- "late"


# Combine the objects using named arguments
emm_predictions <- rbind(early = emm_predictions_early, 
                      late  = emm_predictions_late)

summary(emm_predictions, type = "response") %>%
  as.data.frame()%>%
  # mutate(eventonsettimes=c("early","early","late","late"))%>%
  kbl(escape = F, digits=4,
      caption = paste0("Marginal Mean predicted Probabilities in early and late time window"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))

# Now, set up the contrast to compare early vs. late for each condition.
# This will perform a pairwise comparison for each event type.
# do everything on the response scale, i.e. probability scale
emm_response_scale <- regrid(emm_predictions)

# Now perform the contrasts on the regridded object
contrast_tdiffs <- contrast(
  emm_response_scale,
  method = "pairwise",
  by = "eventtype2",
  adjust= "holm"
)


# View the results on the probability scale
summary(contrast_tdiffs) %>%
  as.data.frame()%>%
  mutate(contrast = c("early vs late"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  mutate(`CohensD1` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = df.residual(model_gam_interaction), method = "identity"))$effect.size
  )%>%
  mutate(`CohensD2` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = sum(model_gam_interaction$edf), method = "identity"))$effect.size
  )%>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of hit rates between early and late time window | Holm corrected"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))



# Now perform the contrasts on the regridded object
contrast_condiffs <- contrast(
  emm_response_scale,
  method = "pairwise",
  by = "eventonsettimes",
  adjust= "holm"
)


# View the results on the probability scale
summary(contrast_condiffs) %>%
  as.data.frame()%>%
  mutate(eventonsettimes = c("early","late"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  mutate(`CohensD1` =
           summary(eff_size(contrast_condiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = df.residual(model_gam_interaction), method = "identity"))$effect.size
  )%>%
  mutate(`CohensD2` =
           summary(eff_size(contrast_condiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = sum(model_gam_interaction$edf), method = "identity"))$effect.size
  )%>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of hit rates between conditions | Holm corrected"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))


emm_predictions_plot <- emmeans(
  model_gam_interaction,
  ~ eventtype2 | eventonsettimes,
  at = list(eventonsettimes = c(mean(early_window), mean(late_window)))
) %>%
  regrid()%>%
  as.data.frame()




# Generate a data frame of predictions for the eventonsettimes effect
predictions_df <- ggpredict(
  model_gam_interaction,
  terms = c("eventonsettimes","eventtype2")
)


 
plotplot <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         eventtype2 = as.factor(eventtype2))%>%
  ggplot(., aes(x = eventonsettimes, y = as.numeric(hit), color = eventtype2)) +
  # Original data points
  stat_summary(fun = mean, geom = "point", alpha = 0.4, bins=length(unique(DATA_Trials$eventonsettimes))) +
   # Add the GAM predictions as a line and a ribbon
  # geom_ribbon(data = predictions_df, aes(x = x, y = predicted,ymin = conf.low, ymax = conf.high, fill = group),
  #             alpha = 0.4, , colour = NA) +
  geom_ribbon(data = predictions_df, aes(x = x, y = predicted,fill = group, ymin = conf.low, ymax = conf.high, group=group),
              alpha = 0.2, , colour = NA) +
  # geom_ribbon(data = predictions_df, aes(x = x, y = predicted,ymin = conf.low, ymax = conf.high), fill = "steelblue", alpha = 0.2) +
  geom_line(data = predictions_df, aes(x = x, y = predicted, color = group), size = 1)+
    # Labels and theme
  labs(x = "Event Onset Time (s)", y = "Probability of Hits",
       title = "Probability of Hits | GAM model", subtitle = "+ predicted hit rates for windows of interest") +
  geom_point(data = emm_predictions_plot, aes(y = prob, group = eventtype2), size = 2, stroke = 1) +
  geom_errorbar(data = emm_predictions_plot, aes(y = prob,ymin = lower.CL, ymax = upper.CL), width = 0.1, size = 1) +
  scale_fill_manual( values=c("#198a83", "#ff8504")) +
  scale_color_manual( values=c("#198a83", "#ff8504")) +
  # Add vertical lines for the early time window
  annotate("segment", x = early_window[1], xend = rev(early_window)[1], y = 0.3, yend = 0.3, color = "grey60", size = 1.5) +
  annotate("segment", x = late_window[1], xend = rev(late_window)[1], y = 0.3, yend = 0.3, color = "grey60", size = 1.5) +
  # theme_minimal()
  theme_light()


savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", height = 3,width = 4.9,  unit = "in")
savefig %<>%
  fill_panel(plotplot, column = 1, row = 1)
savefig
# save_multi_panel_figure(savefig, "figures/Behavior_ProbHits.pdf")

```

<br> 

### Reaction Times

#### Statistical test for time-resolved analysis of reaction times
*Do reaction times for hits for primed and non-primed events differ (differently) across time?*

full model: RT ~ eventtype2 * eventonsettimes + (1 | participant)

reduced model: RT ~ eventtype2 + eventonsettimes + (1 | participant)

<br> 
```{r stat_RTxpresentationtime_distr,  warning = FALSE, fig.height=4, fig.width=6}
# Model with smoothers by condition
# with time x cue interaction
# Model with smoothers by condition
model_gam_interaction <- DATA_Trials %>%
  filter(response=='hit')%>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(RT ~ eventtype2 + s(eventonsettimes, by = eventtype2, bs = "cs") + participant, 
                 data = .)
# coef(model_gam_interaction)
# summary(model_gam_interaction)

# no interaction only main effects
model_gam_TwoMain <- DATA_Trials %>%
filter(response=='hit')%>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(RT ~ eventtype2 + s(eventonsettimes, bs = "cs") + participant, 
                 data = .)

# no effect of time only CUE
model_gam_OneMain <- DATA_Trials %>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(RT ~ eventtype2 +  participant, 
                 data = .)

# no effect of time
model_gam_null <- DATA_Trials %>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(RT ~ participant, 
                 data = .)
  
plot(model_gam_interaction, pages = 1)
# anova(model_gam_interaction)


# based on 'marginal effects'
# plot_predictions(model_gam_interaction, condition = c('eventonsettimes', 'eventtype2'),
#                  type = 'response') +
#       labs(y = "Predicted probability of a hit",
#            title = "Average smooth effect of eventonsettimes | GAM",
#            subtitle = "Per priming condition")


anova(model_gam_null,model_gam_OneMain,model_gam_TwoMain,model_gam_interaction) %>%
  tidy()%>%
  kbl(escape = F, digits=3,
      caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:"))%>%
  # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>",
  #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = 
                                                                   "hold_position"))


AIC(model_gam_null, model_gam_OneMain, model_gam_TwoMain, model_gam_interaction) %>%
  # tidy()%>%
  kbl(escape = F, digits=3,
      caption = paste0("AIC | comparing full and reduced model | GAM:"))%>%
  # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>",
  #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options =
                                                                   "hold_position"))
```

<br> 


##### follow up: directly test time windows | RTs
*Do RTs for primed and non-primed events differ (differently) across time?*

full model: RT ~ eventtype2 * eventonsettimes + (1 | participant)

Test rationale: do these models differ significantly and does the interaction term thus allow for a better explanation of the data.

Extract predicted probability of RTs for early and late window and contrast them

<br> 
```{r stat_RTxpresentationtime_distr_contrasts, warning = FALSE, fig.height=3.5, fig.width=4.9}
# Full model with interaction
# Model with smoothers by condition
model_gam_interaction <- DATA_Trials %>%
  filter(response=='hit')%>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(RT ~ eventtype2 + s(eventonsettimes, by = eventtype2, bs = "cs") + participant, 
                 data = .)

# alternative approach
# Define the time windows
all_timepoints = sort(unique(DATA_Trials$eventonsettimes))
early_window <- all_timepoints[
  (all_timepoints >= min(unique(DATA_Trials$eventonsettimes)) + 0) & (all_timepoints <= min(unique(DATA_Trials$eventonsettimes)) + 0.5)
]
late_window <-all_timepoints[
  (all_timepoints >= max(unique(DATA_Trials$eventonsettimes)) -.5) & (all_timepoints <= max(unique(DATA_Trials$eventonsettimes)) + 0)
]

emm_predictions_early <- emmeans(
  model_gam_interaction,
  ~ eventtype2 ,
  at = list(eventonsettimes = early_window)
)

emm_predictions_late <- emmeans(
  model_gam_interaction,
  ~ eventtype2 ,
  at = list(eventonsettimes = late_window)
)

# 1. Add the identifier column to the grid of each object
emm_predictions_early@grid$eventonsettimes <- "early"
emm_predictions_late@grid$eventonsettimes  <- "late"


# Combine the objects using named arguments
emm_predictions <- rbind(early = emm_predictions_early, 
                      late  = emm_predictions_late)

summary(emm_predictions, type = "response") %>%
  as.data.frame()%>%
  # mutate(eventonsettimes=c("early","early","late","late"))%>%
  kbl(escape = F, digits=4,
      caption = paste0("Marginal Mean predicted RTs in early and late time window"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))

# Now, set up the contrast to compare early vs. late for each condition.
# This will perform a pairwise comparison for each event type.
# do everything on the response scale, i.e. probability scale
emm_response_scale <- regrid(emm_predictions)

# Now perform the contrasts on the regridded object
contrast_tdiffs <- contrast(
  emm_response_scale,
  method = "pairwise",
  by = "eventtype2",
  adjust= "holm"
)


# View the results on the probability scale
summary(contrast_tdiffs) %>%
  as.data.frame()%>%
  mutate(contrast = c("early vs late"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  mutate(`CohensD1` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = df.residual(model_gam_interaction), method = "identity"))$effect.size
  )%>%
  mutate(`CohensD2` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = sum(model_gam_interaction$edf), method = "identity"))$effect.size
  )%>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of RTs between early and late time window | Holm corrected"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))


# Now perform the contrasts on the regridded object
contrast_condiffs <- contrast(
  emm_response_scale,
  method = "pairwise",
  by = "eventonsettimes",
  adjust= "holm"
)


# View the results on the probability scale
summary(contrast_condiffs) %>%
  as.data.frame()%>%
  mutate(eventonsettimes = c("early","late"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  mutate(`CohensD1` =
           summary(eff_size(contrast_condiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = df.residual(model_gam_interaction), method = "identity"))$effect.size
  )%>%
  mutate(`CohensD2` =
           summary(eff_size(contrast_condiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = sum(model_gam_interaction$edf), method = "identity"))$effect.size
  )%>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of RTs between conditions | Holm corrected"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))


emm_predictions_plot <- emmeans(
  model_gam_interaction,
  ~ eventtype2 | eventonsettimes,
  at = list(eventonsettimes = c(mean(early_window), mean(late_window)))
) %>%
  regrid()%>%
  as.data.frame()



# Generate a data frame of predictions for the eventonsettimes effect
predictions_df <- ggpredict(
  model_gam_interaction,
  terms = c("eventonsettimes","eventtype2")
)

plotplot <- DATA_Trials %>%
  filter(response=='hit')%>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  ggplot(., aes(x = eventonsettimes, y = as.numeric(RT), color = eventtype2)) +
  # Original data points
  stat_summary(fun = mean, geom = "point", alpha = 0.4, bins=length(unique(DATA_Trials$eventonsettimes))) +
  
  # Add the GAM predictions as a line and a ribbon
  # geom_ribbon(data = predictions_df, aes(x = x, y = predicted,ymin = conf.low, ymax = conf.high, fill = group),
  #             alpha = 0.4, , colour = NA) +
  geom_ribbon(data = predictions_df, aes(x = x, y = predicted,fill = group, ymin = conf.low, ymax = conf.high, group=group),
              alpha = 0.2, , colour = NA) +
  # geom_ribbon(data = predictions_df, aes(x = x, y = predicted,ymin = conf.low, ymax = conf.high), fill = "steelblue", alpha = 0.2) +
  geom_line(data = predictions_df, aes(x = x, y = predicted, color = group), size = 1)+
    # Labels and theme
  labs(x = "Event Onset Time (s)", y = "RT",
       title = "RT | GAM model", subtitle = "+ predicted RT for windows of interest") +
  geom_point(data = emm_predictions_plot, aes(y = emmean, group = eventtype2), size = 2,  stroke = 1) +
  geom_errorbar(data = emm_predictions_plot, aes(y = emmean,ymin = lower.CL, ymax = upper.CL), width = 0.1, size = 1) +
  scale_fill_manual( values=c("#198a83", "#ff8504")) +
  scale_color_manual( values=c("#198a83", "#ff8504")) +
  # Add vertical lines for the early time window
  annotate("segment", x = early_window[1], xend = rev(early_window)[1], y = 500, yend = 500, color = "grey60", size = 1.5) +
  annotate("segment", x = late_window[1], xend = rev(late_window)[1], y = 500, yend = 500, color = "grey60", size = 1.5) +
  # theme_minimal()+
  theme_light() +
  guides(fill="none")

savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", height = 3,width = 4.9,  unit = "in")
savefig %<>%
  fill_panel(plotplot, column = 1, row = 1)
savefig
# save_multi_panel_figure(savefig, "figures/Behavior_ProbRT.pdf")

```
<br> 


<br> 

### False Alarm Rates

#### Display of smoothed False Alarm Rates across time
What are the distributions of false alarm rates across the trial?

<br> 

```{r disp_FA_X_TIME, fig.height=4, fig.width=7, warning = FALSE}
DATA_FA <- DataIn1%>%
  filter(eventtype=="distractor")%>%
  mutate(FA = ifelse(response == "FA_proper", 1, 0))

ggplot(DATA_FA,aes(x = eventonsettimes, y = FA)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "loess", se = TRUE, span = 0.3) +    # kernel smoothed average
  labs(
    title = "Smoothed FA probabilities",
    x = "eventonsettimes (s)",
    y = "FA probability"
  ) +
  theme_minimal()

ggplot(DATA_FA, aes(x = eventonsettimes, y = FA)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), se = TRUE) +
  labs(x = "Event Onset Time", y = "FA probability", 
       title = "GAM Smoothed Probability of FAs") +
  theme_minimal()



```
<br> 

#### Statistical test for time-resolved analysis of FArate
*Does the probability of FAs differ across time?*

full model: FA ~  eventonsettimes + (1 | participant)
reduced model: FA ~  1 + (1 | participant)

Test rationale: do these models differ significantly and does the interaction term thus allow for a better explanation of the data.

<br> 
```{r stat_FA_X_TIME_test, warning = FALSE, fig.height=4, fig.width=6}


# Full model with factor
GLMHit_full <- DataIn1%>%
  filter(eventtype=="distractor")%>%
  mutate(response = as.factor(response)) %>%
  glmer(response ~ eventonsettimes + (1 | participant),data=., family = binomial)

# Reduced model intercept model
GLMHit_reduced <- DataIn1%>%
  filter(eventtype=="distractor")%>%
  mutate(response = as.factor(response)) %>%
  glmer(response ~ 1 + (1 | participant),data=., family = binomial)

# Likelihood ratio test
anova( GLMHit_reduced, GLMHit_full, test = "Chisq") %>%
  as.data.frame() %>%
  kable(escape = F, digits=3,
        caption = paste0("Likelihood ratio test via anova | comparing full and reduced model:<br>", 
                    deparse(formula(GLMHit_full))))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))

# Display the summary of the full model + Wald test
GLMHit_full %>%
  tidy() %>%
  kable(escape = F, digits=3,
        caption = paste0("Likelihood ratio test via anova | comparing full and reduced model:<br>", 
                    deparse(formula(GLMHit_full)))) %>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8))


# Model with smoothers by condition
DATA_FA <- DataIn1%>%
  filter(eventtype=="distractor")%>%
  mutate(FA = ifelse(response == "FA_proper", 1, 0))



model_gam_factor <- DATA_FA%>%
  gam(FA ~ s(eventonsettimes, bs = "cs") + participant, 
                 data = ., family = binomial)

model_gam_null <-DATA_FA%>%
  gam(FA ~ participant, 
                 data = ., family = binomial)

plot(model_gam_factor, pages = 1)
# anova(model_gam_interaction)


# plot_predictions(model_gam_factor, condition = 'eventonsettimes',
#                  type = 'response') +
#       labs(y = "Predicted probability of a FA",
#            title = "Average smooth effect of eventonsettimes | GAM")


# Likelihood ratio test
anova( model_gam_null, model_gam_factor, test = "Chisq") %>%
  tidy()%>%
  # as.data.frame() %>%
  # mutate(model = c('reduced model','full model')) %>%
  # dplyr::select(model, `Resid. Df`:`Pr(>Chi)`)%>%
  kbl(escape = F, digits=3,
      caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))



ggplot(DATA_FA, aes(x = eventonsettimes, y = FA)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"),family = binomial,  se = TRUE) +
  labs(x = "Event Onset Time", y = "FA probability", 
       title = "GAM Smoothed Probability of FAs") +
  theme_minimal()


# Generate a data frame of predictions for the eventonsettimes effect
predictions_df <- ggpredict(
  model_gam_factor,
  terms = "eventonsettimes"
)



```

<br> 

##### follow up: directly test time windows | FAs
*Do FAs for primed and non-primed events differ (differently) across time?*

full model: FA ~ eventtype2 * eventonsettimes + (1 | participant)

Test rationale: do these models differ significantly and does the interaction term thus allow for a better explanation of the data.

Extract predicted probability of FAs for early and late window and contrast them

<br> 
```{r stat_FAxpresentationtime_distr_contrasts, warning = FALSE, fig.height=4, fig.width=7}
# Full model with interaction
# Model with smoothers by condition
model_gam_interaction <- DataIn1%>%
  filter(eventtype=="distractor")%>%
  mutate(FA = ifelse(response == "FA_proper", 1, 0))%>%
  gam(FA ~ s(eventonsettimes, bs = "cs") + participant, 
                 data = ., family = binomial)



# # based on gam interactions
# # Define the time windows
# early_window <- c(0, 0.5) + min(DataIn1$eventonsettimes)
# late_window <- c(-0.5, 0) + max(DataIn1$eventonsettimes)
# 
# # Calculate the average predicted hit probability for each time window
# # and for each event type (attended vs. unattended).
# # We'll use the 'at' argument to specify the time points of interest.
# emm_predictions <- emmeans(
#   model_gam_interaction,
#   ~ eventonsettimes,
#   at = list(eventonsettimes = c(mean(early_window), mean(late_window)))
# )


# alternative approach
# Define the time windows
all_timepoints = sort(unique(DATA_Trials$eventonsettimes))
early_window <- all_timepoints[
  (all_timepoints >= min(unique(DATA_Trials$eventonsettimes)) + 0) & (all_timepoints <= min(unique(DATA_Trials$eventonsettimes)) + 0.5)
]
late_window <-all_timepoints[
  (all_timepoints >= max(unique(DATA_Trials$eventonsettimes)) -.5) & (all_timepoints <= max(unique(DATA_Trials$eventonsettimes)) + 0)
]

emm_predictions_early <- emmeans(
  model_gam_interaction,
  ~ 1 ,
  at = list(eventonsettimes = early_window)
)

emm_predictions_late <- emmeans(
  model_gam_interaction,
  ~ 1 ,
  at = list(eventonsettimes = late_window)
)

# 1. Add the identifier column to the grid of each object
emm_predictions_early@grid$eventonsettimes <- "early"
emm_predictions_late@grid$eventonsettimes  <- "late"


# Combine the objects using named arguments
emm_predictions <- rbind(early = emm_predictions_early, 
                      late  = emm_predictions_late)




summary(emm_predictions, type = "response") %>%
  as.data.frame()%>%
  # mutate(eventonsettimes=c("early","late"))%>%
  kbl(escape = F, digits=4,
      caption = paste0("Marginal Mean predicted RTs in early and late time window"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))


# Now, set up the contrast to compare early vs. late for each condition.
# This will perform a pairwise comparison for each event type.
# do everything on the response scale, i.e. probability scale
emm_response_scale <- regrid(emm_predictions)

# Now perform the contrasts on the regridded object
contrast_tdiffs <- contrast(
  emm_response_scale,
  method = "pairwise"
)


# View the results on the probability scale
summary(contrast_tdiffs) %>%
  as.data.frame()%>%
  mutate(contrast = c("early vs late"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  mutate(`CohensD1` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = df.residual(model_gam_interaction), method = "identity"))$effect.size
  )%>%
  mutate(`CohensD2` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = sum(model_gam_interaction$edf), method = "identity"))$effect.size
  )%>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of FAs between early and late time window"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))


# Now perform the contrasts on the regridded object
contrast_condiffs <- contrast(
  emm_response_scale,
  method = "pairwise"
)


emm_predictions_plot <- emmeans(
  model_gam_interaction,
  ~ eventonsettimes,
  at = list(eventonsettimes = c(mean(early_window), mean(late_window)))
) %>%
  regrid()%>%
  as.data.frame()

# Base plot with original data and GAM smoothed line
 DataIn1%>%
  filter(eventtype=="distractor")%>%
  mutate(FA = ifelse(response == "FA_proper", 1, 0))%>%
  ggplot(., aes(x = eventonsettimes, y = as.numeric(FA))) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +
  # stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.02, alpha = 0.4) + # Optional: add standard error bars
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), se = TRUE) +
  labs(x = "Event Onset Time (s)", y = "Probability of Hits",
       title = "Probability of Hits | GAM model", subtitle = "+ predicted hit rates for windows of interest") +
  geom_point(data = emm_predictions_plot, aes(y = prob), size = 3, shape = 23, stroke = 1.5) +
  geom_errorbar(data = emm_predictions_plot, aes(y = prob,ymin = lower.CL, ymax = upper.CL), width = 0.1, size = 1) +
  # Add vertical lines for the early time window
  annotate("segment", x = early_window[1], xend = early_window[2], y = -0.01, yend = -0.01, color = "grey60", size = 1.5) +
  annotate("segment", x = late_window[1], xend = late_window[2], y = -0.01, yend = -0.01, color = "grey60", size = 1.5) +
  theme_minimal()


# Generate a data frame of predictions for the eventonsettimes effect
predictions_df <- ggpredict(
  model_gam_factor,
  terms = "eventonsettimes"
)

ggplot(DATA_FA, aes(x = eventonsettimes, y = FA)) +
  # Original data points
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +
  # Add the GAM predictions as a line and a ribbon
  geom_line(data = predictions_df, aes(x = x, y = predicted), color = "grey60", size = 1) +
  geom_ribbon(data = predictions_df, aes(x = x, y = predicted,ymin = conf.low, ymax = conf.high), fill = "grey60", alpha = 0.2) +
  # Labels and theme
  labs(x = "Event Onset Time (s)", y = "Probability of FAs",
       title = "Probability of FAs | GAM model", subtitle = "+ predicted FA rates for windows of interest") +
  geom_point(data = emm_predictions_plot, aes(y = prob), size = 2,  stroke =1) +
  geom_errorbar(data = emm_predictions_plot, aes(y = prob,ymin = lower.CL, ymax = upper.CL), width = 0.1, size = 1) +
  # Add vertical lines for the early time window
  annotate("segment", x = early_window[1], xend = rev(early_window)[1], y = -0.01, yend = -0.01, color = "grey60", size = 1.5) +
  annotate("segment", x = late_window[1], xend = rev(late_window)[1], y = -0.01, yend = -0.01, color = "grey60", size = 1.5) +
  theme_light()




```

<br> 

### Control Analysis: Are there differences in the distributions of event times for both cueing conditons?

Idea: can we really compare primed and non-primed events? Are their distributions of presentation times comparable?

Rationale: Can the eventtype be predicted by the presentation time? If presentation time was comparable, it shouldn't be a predictor.

eventtype2 ~ eventonsettimes + (1 | participant)

<br> 
```{r stat_hitxpresentationtime_distr_control, fig.show = "hide", warning = FALSE, fig.height=4, fig.width=10}
GLMEventType_full <- DATA_Trials %>%
  mutate(eventtype2 = as.factor(eventtype2)) %>%
  glmer(eventtype2 ~ eventonsettimes + (1 | participant),data=., family = binomial)


# Display the summary of the full model + Wald test
# summary(GLMEventType_full)

# Display the summary of the full model + Wald test
GLMEventType_full %>%
  tidy() %>%
  kable(escape = F, digits=3,
        caption = paste0("Likelihood ratio test via anova | display of full model:<br>", 
                    deparse(formula(GLMEventType_full)))) %>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8))

```

<br> 

### Control Analysis: Is the sequential modulation simply capturing a sequential strategy in shifting attention by the to be learned cue?

Idea: Can the sequential modulation be solely attributed to the functional classification primed vs non-primed? Or could an alternative account state that the sequential modulation is capturing a sequential strategy in shifting attention by the to be learned cue? Does it even make a difference? If it is learning related, maybe the effect is time-on-task dependent and initially higher than later on.


Rationale: Do a time-on-task resolved analysis

eventtype2 ~ eventonsettimes + (1 | participant)

<br> 
```{r stat_control_learning, fig.height=4, fig.width=8, warning = FALSE}
data_bin <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0))%>%
  group_by(participant)%>%
  # mutate(time_bin = cut(trialnumber,
  #                       breaks = seq(0, max(trialnumber), by = 2),
  #                       labels = FALSE)) %>%
  mutate(time_bin = as.factor(ntile(trialnumber, 2))) %>% 
  ungroup()



ggplot(data_bin, aes(x = eventonsettimes, y = hit, color = eventtype2)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "loess", span = 0.3, se = TRUE) +
  facet_grid(.~time_bin)+
  labs(x = "Event Onset Time", y = "Probability of hits", 
       title = "Smoothed Probability of hits | single sub data") +
  theme_minimal()

# ggplot(data_avg, aes(x = eventonsettimes, y = hit_rate, color = eventtype2)) +
#   # geom_line(alpha = 0.4) +
#   stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
#   geom_smooth(method = "loess", se = TRUE, span = 0.3) +
#   labs(x = "Event Onset Time", y = "Hit Rate", 
#        title = "Smoothed Hit Rates Over Time by Condition | averaged across subjects hit rates") +
#   theme_minimal()


ggplot(data_bin, aes(x = eventonsettimes, y = hit, color = eventtype2)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), se = TRUE) +
  facet_grid(.~time_bin)+
  labs(x = "Event Onset Time", y = "Probability of hits", 
       title = "GAM Smoothed Probability of Hits by Condition") +
  theme_minimal()



data_bin %>%
  filter(response=='hit')%>%
  ggplot(aes(x = eventonsettimes, y = RT, color = eventtype2)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "loess", se = TRUE, span = 0.3) +    # kernel smoothed average
  facet_grid(.~time_bin)+
  labs(
    title = "Smoothed Average Reaction Times",
    x = "eventonsettimes (s)",
    y = "Reaction Time (ms)",
    color = "Condition"
  ) +
  theme_minimal()

data_bin %>%
  filter(response=='hit')%>%
  ggplot(aes(x = eventonsettimes, y = RT, color = eventtype2)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), se = TRUE) +    # kernel smoothed average
  facet_grid(.~time_bin)+
  labs(
    title = "GAM Smoothed Average Reaction Times",
    x = "eventonsettimes (s)",
    y = "Reaction Time (ms)",
    color = "Condition"
  ) +
  theme_minimal()



# hit rates
# Model with smoothers by condition
model_gam_null <- data_bin%>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(hit ~ eventtype2 + s(eventonsettimes, by = eventtype2, bs = "cs") + participant, 
                 data = ., family = binomial)
# coef(model_gam_interaction)



model_gam_timeOntask <- data_bin%>%
  mutate(
    eventtype2 = as.factor(eventtype2),
    time_bin   = as.factor(time_bin),
    eventtype2_timebin = interaction(eventtype2, time_bin, drop = TRUE)
  ) %>%
  gam(hit ~ eventtype2*time_bin + 
        s(eventonsettimes, by = eventtype2_timebin, bs = "cs") + 
        participant, 
                 data = ., family = binomial)

plot(model_gam_timeOntask, pages = 1)
# anova(model_gam_interaction)


# plot_predictions(model_gam_timeOntask, condition = c('eventonsettimes', 'eventtype2_timebin'),
#                  type = 'response') +
#       labs(y = "Predicted probability of a hit",
#            title = "Average smooth effect of eventonsettimes | GAM",
#            subtitle = "Per priming condition")


# Likelihood ratio test
anova( model_gam_null, model_gam_timeOntask, test = "Chisq") %>%
  tidy()%>%
  # as.data.frame() %>%
  # mutate(model = c('reduced model','full model')) %>%
  # dplyr::select(model, `Resid. Df`:`Pr(>Chi)`)%>%
  kbl(escape = F, digits=3,
      caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))




# reaction times
# Model with smoothers by condition
model_gam_null <- data_bin%>%
  filter(response=='hit')%>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(RT ~ eventtype2 + s(eventonsettimes, by = eventtype2, bs = "cs") + participant, 
                 data = .)
# coef(model_gam_interaction)



model_gam_timeOntask <- data_bin%>%
  filter(response=='hit')%>%
  mutate(
    eventtype2 = as.factor(eventtype2),
    time_bin   = as.factor(time_bin),
    eventtype2_timebin = interaction(eventtype2, time_bin, drop = TRUE)
  ) %>%
  gam(RT ~ eventtype2*time_bin + 
        s(eventonsettimes, by = eventtype2_timebin, bs = "cs") + 
        participant, 
                 data = .)


plot(model_gam_timeOntask, pages = 1)
# anova(model_gam_interaction)


# plot_predictions(model_gam_timeOntask, condition = c('eventonsettimes', 'eventtype2_timebin'),
#                  type = 'response') +
#       labs(y = "RT",
#            title = "Average smooth effect of eventonsettimes | GAM",
#            subtitle = "Per priming condition")


# Likelihood ratio test
anova( model_gam_null, model_gam_timeOntask, test = "Chisq") %>%
  tidy()%>%
  # as.data.frame() %>%
  # mutate(model = c('reduced model','full model')) %>%
  # dplyr::select(model, `Resid. Df`:`Pr(>Chi)`)%>%
  kbl(escape = F, digits=3,
      caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))

```