---
title: "behavioral analysis"
author: "Christopher"
date: "16 6 2025"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 5
    theme: flatly
    df_print: paged
---



```{r load_package, message = FALSE, warning = FALSE}
rm(list=ls()) # clear environment

library(lme4)
library(readxl)
library(tidyverse)
# library(data.table)
# library(afex)
# library(broom)
# library(apa)
library(kableExtra)
# library(lmerTest)
# library(pbkrtest)
# library(effects)
# library(visreg)
# library(sjPlot)
library(broom.mixed)
# library(pander)
# library(mediation)
# library(multcomp)
# library(multcompView)
# library(magrittr)
library(multipanelfigure)
library(ggbeeswarm)
# library(lsmeans)
library(BayesFactor)
library(ggpubr)
# library(gpairs)
# library(DescTools)
library(cowplot)
# library(ggpol)
# library(psych)
library(ggdist)
# library(dplyr)
# 
library(mgcv)
library(tidyverse)       # Tidy and flexible data manipulation
# library(marginaleffects) # Compute conditional and marginal effects
# library(patchwork)       # Combining ggplot objects
# library(gratia)
# 
library(ggeffects)
# 
library(report)
library(emmeans)

library(forecast)


source('C:/Dropboxdata/Dropbox/work/R-statistics/general_functions/RainCloudPlots/tutorial_R/R_rainclouds.R')





# broad cluster
# CSD based normalized to pre-cue baseline in %
# 
DATAPath1 = "data_in/behavior_events.csv"
DATAPath2 = "data_in/behavior_FAs.csv"
DATAPath3 = "data_in/Gabor_TimeCourses_01-07-2026_09-42.csv"





options(scipen=1, digits=5)
```
<style type="text/css">
.main-container {
  max-width: 1800px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>


## Prepare data  
<br>  

1.  Read in data: 
    + `r DATAPath1`
    + `r DATAPath1`
    
<br>  

```{r load_data,results = "hide", fig.show = "hide", warning = FALSE}
sub_sample = as.factor(c(5)) # no ssvep people discarded


# read in data
DataIn1 <- read_csv(DATAPath1) %>%
  filter(!participant %in% sub_sample)
DataIn2 <- read_csv(DATAPath2)%>%
  filter(!subject %in% sub_sample)
DataIn3 <- read_delim(DATAPath3, delim =";")%>%
  filter(!participant %in% sub_sample)%>%
  mutate(condition = as.factor(condition)) %>%
  mutate(time = time/1000) # data should be in s
head(DataIn1)
str(DataIn1)
head(DataIn2)
str(DataIn2)
head(DataIn3)
str(DataIn3)

#


```





1.  Prepare new dataset(s)...
<br>  

```{r prepare_data1, results = "hide", fig.show = "hide", warning = FALSE}
DATA_Trials <- DataIn1%>%
  filter(eventtype=="target")


```

## How to cite the packages?

<br> 
```{r cite data, results = "hide", fig.show = "hide", warning = FALSE}
session <- sessionInfo()
r <- report(session)
r
summary(r)
as.data.frame(r)
summary(as.data.frame(r))

```

## Illustrate behavioral data

<br>  

```{r plot_data_avg, results = "hide",  fig.height=6, fig.width=7, warning = FALSE}

dat2plot1 <- DATA_Trials %>%
  group_by(participant, eventtype2)%>%
  dplyr::summarise(
    RT_Mean=mean(RT,na.rm=TRUE),
    RT_STD=sd(RT,na.rm=TRUE),
    Hitrate=(sum(response=='hit')/n())*100,
    Hitrate_adj1 = ifelse((sum(response=='hit')/n())>0,
                         min(c(sum(response=='hit')/n(), 1-(.5/n()))),
                         .5/n()), # according to Stanislaw H, Todorov N, Behav Res Meth (1999) 31, 137-149, "1/2N rule" or
    Hitrate_adj2 = (sum(response=='hit')+0.5)/(n()+1))%>% # loglinear approach (Hautus, 1995)
    ungroup()
dat2plot1a = DataIn1%>%
  filter(eventtype=="distractor") %>%
  group_by(participant, eventtype2)%>%
  dplyr::summarise(
    FArate=(sum(response=='FA_proper')/n())*100,
    FArate_adj1 = ifelse((sum(response=='FA_proper')/n())>0,
                         min(c(sum(response=='FA_proper')/n(), 1-(.5/n()))),
                         .5/n()), # according to Stanislaw H, Todorov N, Behav Res Meth (1999) 31, 137-149, "1/2N rule" or
    FArate_adj2 = (sum(response=='FA_proper')+0.5)/(n()+1)) %>% # loglinear approach (Hautus, 1995))
  ungroup()
dat2plot <- merge(dat2plot1,dat2plot1a,by = "participant") %>% # adjust for HR and FR of 0 or 1
  mutate(
    d_prime = qnorm(Hitrate_adj1)-qnorm(FArate_adj1),
    c = -0.5*(qnorm(Hitrate_adj1)+qnorm(FArate_adj1))
  ) %>%
  dplyr::rename(eventtype2=eventtype2.x)




dat2plot2 <- DataIn2 %>%
  mutate(`FA_all`= FA + `FA_proper`)%>%
  # dplyr::select(-FA)%>%
  dplyr::select(-FA_all)%>%
  pivot_longer(c(-subject ),names_to = "FA_type", values_to = "number")%>%
  filter(FA_type=='FA_proper' | FA_type=='FA')

theme_set(theme_bw())
plot1 <- ggplot(dat2plot, aes(x = 1, y = RT_Mean)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  ylab(expression(paste("reaction time in ms")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

plot2 <- ggplot(dat2plot, aes(x = 1, y = Hitrate)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange"))
  ylab(expression(paste("hit rate in %")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

plot3 <- ggplot(dat2plot2, aes(x = FA_type, y = number, color = FA_type)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=3.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  scale_fill_manual("FA type", values=c("grey40", "black"),labels = c("to other","to distractor")) +
  scale_color_manual("FA type", values=c("grey40", "black"),labels = c("to other","to distractor")) +
  ylab(expression(paste("number of False Alarms")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

plot4 <- ggplot(dat2plot, aes(x = 1, y = d_prime)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange"))
  ylab(expression(paste("dprime")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())


# cowplot::plot_grid(plot1, plot2, plot3, ncol = 3,labels = "AUTO")
plotttitle <- ggdraw() + draw_label("behavioral measures", fontface='bold')
plotplots <- ggarrange(plot1, plot2, plot3, plot4, ncol = 4, align = "h", widths=c(2,2,5,2))
ggarrange(plotttitle, plotplots, nrow = 2, heights=c(0.2, 1))




## for publication
theme_set(theme_bw())
plot1 <- ggplot(dat2plot, aes(x = eventtype2, y = RT_Mean, color =eventtype2,)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  ylab(expression(paste("reaction time in ms")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_y_continuous(limits=c(400, 775))

plot1b <- ggplot(dat2plot, aes(x = 1, y = RT_Mean, color =eventtype2, fill = eventtype2)) +
  geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.0, trim = FALSE, alpha = 0.8)+
  # scale_fill_manual(values=c("grey40"))+
  scale_fill_brewer(palette = "Dark2")+
  # scale_color_manual(values=c("grey40"))+
  scale_color_brewer(palette = "Dark2")+
  theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
         axis.ticks.y=element_blank())+
  theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
         axis.ticks.x=element_blank())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_y_continuous(limits=c(400, 775))

plot2 <- ggplot(dat2plot, aes(x = eventtype2, y = Hitrate, color =eventtype2)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
    # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange"))
  ylab(expression(paste("hit rate in %")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_y_continuous(limits=c(30, 100))

plot2b <- ggplot(dat2plot, aes(x = 1, y = Hitrate,  color =eventtype2, fill = eventtype2)) +
  geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.0, trim = FALSE, alpha = 0.8)+
  # scale_fill_manual(values=c("grey40"))+
  scale_fill_brewer(palette = "Dark2")+
  # scale_color_manual(values=c("grey40"))+
  scale_color_brewer(palette = "Dark2")+
  theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
         axis.ticks.y=element_blank())+
  theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
         axis.ticks.x=element_blank())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_y_continuous(limits=c(30, 100))

  

# cowplot::plot_grid(plot1, plot2, plot3, ncol = 3,labels = "AUTO")
plotttitle <- ggdraw() + draw_label("behavioral measures", fontface='bold')
plotplots <- ggarrange(plot1, plot1b,plot2, plot2b, ncol = 4,common.legend = TRUE,align = "h", widths=c(3,2,3,2))

savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", width = 5, height = 6, unit = "in")
saveplot <- ggarrange(plotttitle, plotplots, nrow = 2, heights=c(0.2, 1))

savefig %<>%
  fill_panel(saveplot, column = 1, row = 1)
savefig
# save_multi_panel_figure(savefig, "hits_reactiontimes.eps")


## for publication
theme_set(theme_bw())
plot1 <- ggplot(dat2plot, aes(x = eventtype2, y = RT_Mean, color =eventtype2)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  ylab(expression(paste("reaction time in ms")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_y_continuous(limits=c(400, 775))

plot1b <- ggplot(dat2plot, aes(x = 1, y = RT_Mean, color =eventtype2, fill = eventtype2)) +
  geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.0, trim = FALSE, alpha = 0.8)+
  # scale_fill_manual(values=c("grey40"))+
  scale_fill_brewer(palette = "Dark2")+
  # scale_color_manual(values=c("grey40"))+
  scale_color_brewer(palette = "Dark2")+
  theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
         axis.ticks.y=element_blank())+
  theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
         axis.ticks.x=element_blank())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_y_continuous(limits=c(400, 775))

plot2 <- ggplot(dat2plot, aes(x = eventtype2, y = d_prime, color =eventtype2)) +
  # geom_hline(yintercept=0, show.legend = FALSE) +
  geom_beeswarm(cex=4.5, size = 5,alpha=0.4)+
  stat_summary(fun.y=mean, aes(ymin=..y.., ymax=..y..), geom='errorbar', width=0.9, size=1.5) +
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  # scale_fill_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange")) +
  # scale_color_manual("attended color", values=c("darkcyan", "darkgreen", "darkorange"))
  ylab(expression(paste("dprime")))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_y_continuous(limits=c(0.0, 3.8))

plot2b <- ggplot(dat2plot, aes(x = 1, y = d_prime, color =eventtype2, fill = eventtype2)) +
  geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.0, trim = FALSE, alpha = 0.8)+
  # scale_fill_manual(values=c("grey40"))+
  scale_fill_brewer(palette = "Dark2")+
  # scale_color_manual(values=c("grey40"))+
  scale_color_brewer(palette = "Dark2")+
  theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
         axis.ticks.y=element_blank())+
  theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
         axis.ticks.x=element_blank())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_y_continuous(limits=c(0.0, 3.8))

  

# cowplot::plot_grid(plot1, plot2, plot3, ncol = 3,labels = "AUTO")
plotttitle <- ggdraw() + draw_label("behavioral measures", fontface='bold')
plotplots <- ggarrange(plot1, plot1b,plot2, plot2b, ncol = 4,common.legend = TRUE,align = "h", widths=c(3,2,3,2))

savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", width = 5, height = 6, unit = "in")
saveplot <- ggarrange(plotttitle, plotplots, nrow = 2, heights=c(0.2, 1))

savefig %<>%
  fill_panel(saveplot, column = 1, row = 1)
savefig
# save_multi_panel_figure(savefig, "dprime_reactiontimes.eps")
  

```


## Display behavioral data numerically

<br>  

```{r disp_data_avg, warning = FALSE}
dat2plot1 <- DATA_Trials %>%
  group_by(eventtype2, participant )%>%
  dplyr::summarise(
    RT_Mean=mean(RT,na.rm=TRUE),
    RT_STD=sd(RT,na.rm=TRUE),
    Hitrate=(sum(response=='hit')/n())*100,
    Hitrate_adj1 = ifelse((sum(response=='hit')/n())>0,
                         min(c(sum(response=='hit')/n(), 1-(.5/n()))),
                         .5/n()), # according to Stanislaw H, Todorov N, Behav Res Meth (1999) 31, 137-149, "1/2N rule" or
    Hitrate_adj2 = (sum(response=='hit')+0.5)/(n()+1)) # loglinear approach (Hautus, 1995)

dat2plot1a = DataIn1%>%
  filter(eventtype=="distractor") %>%
  group_by(eventtype2, participant)%>%
  dplyr::summarise(
    FArate=(sum(response=='FA_proper')/n())*100,
    FArate_adj1 = ifelse((sum(response=='FA_proper')/n())>0,
                         min(c(sum(response=='FA_proper')/n(), 1-(.5/n()))),
                         .5/n()), # according to Stanislaw H, Todorov N, Behav Res Meth (1999) 31, 137-149, "1/2N rule" or
    FArate_adj2 = (sum(response=='FA_proper')+0.5)/(n()+1)) # loglinear approach (Hautus, 1995))

dat2plot <- merge(dat2plot1,dat2plot1a,by = "participant") %>% # adjust for HR and FR of 0 or 1
  mutate(
    d_prime = qnorm(Hitrate_adj1)-qnorm(FArate_adj1),
    c = -0.5*(qnorm(Hitrate_adj1)+qnorm(FArate_adj1))
  ) %>%
  dplyr::rename(eventtype2=eventtype2.x)

dat2plot  %>%
  group_by(eventtype2)%>%
  dplyr::summarise(N = n(), mean_RT = mean(RT_Mean), std_RT = sd(RT_Mean), mean_Hitrate = mean(Hitrate), std_Hitrate = sd(Hitrate),
            mean_FArate = mean(FArate), std_FArate = sd(FArate), mean_dprime = mean(d_prime), std_dprime = sd(d_prime)) %>%
  kable(escape = F, digits = c(3,3,3,3,3,3,3,3,3,3), caption = c("descriptives of behavioral measures")) %>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8))

dat2plot  %>%
  group_by(participant)%>%
  dplyr::summarise(N = n(), mean_RT = mean(RT_Mean), std_RT = sd(RT_Mean), mean_Hitrate = mean(Hitrate), std_Hitrate = sd(Hitrate),
            mean_FArate = mean(FArate), std_FArate = sd(FArate), mean_dprime = mean(d_prime), std_dprime = sd(d_prime)) %>%
  ungroup()%>%
  dplyr::summarise(N = n(), M_RT = mean(mean_RT), SD_RT = sd(mean_RT), M_Hitrate = mean(mean_Hitrate), SD_Hitrate = sd(mean_Hitrate),
            M_FArate = mean(mean_FArate), SD_FArate = sd(mean_FArate), M_dprime = mean(mean_dprime), SD_dprime = sd(mean_dprime)) %>%
  kable(escape = F, digits = c(3,3,3,3,3,3,3,3,3,3), caption = c("descriptives of behavioral measures")) %>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8))

```



## Statistical analysis

### ANOVA models
### linear mixed models

#### test RT, Hitrate and dprime between primed and non primed targets

**RT** ~ **priming**

<br> 

```{r lme_RT_lum, fig.show = "hide", warning = FALSE, fig.height=4, fig.width=10}
dat2plot2 <- DataIn2 %>%
  mutate(`FA_all`= FA + `FA_proper`)%>%
  # dplyr::select(-FA)%>%
  dplyr::select(-FA_all)%>%
  pivot_longer(c(-subject ),names_to = "FA_type", values_to = "number")%>%
  filter(FA_type=='FA_proper' | FA_type=='FA')


dat2plot1 <- DATA_Trials %>%
  group_by(participant,eventtype2)%>%
  dplyr::summarise(RT_Mean=mean(RT,na.rm=TRUE),RT_STD=sd(RT,na.rm=TRUE),Hitrate=(sum(response=='hit')/n())*100)

dat2plot2 <- DataIn2 %>%
  mutate(`FA_all`= FA + `FA_proper`)%>%
  # dplyr::select(-FA)%>%
  dplyr::select(-FA_all)%>%
  dplyr::rename(participant=subject)

dat2plot_all <- merge(dat2plot1,dat2plot2,by = "participant") %>% # adjust for HR and FR of 0 or 1
  mutate(
    d_prime = qnorm(Hitrate)-qnorm(FA_proper_rate),
    c = -0.5*(qnorm(Hitrate)+qnorm(FA_proper_rate))
  )



# # statisticical analysis ANOVA
# StatsOut <- dat2plot_all %>%
#   aov_ez(id="participant", dv = "RT_Mean", data = ., between = c("colorlum"), include_aov = afex_options("include_aov"))
# 
# StatsOut%>%
#   .$anova %>%
#   mutate(factor = rownames(.))%>%
#   mutate(
#     `Pr(>F)` = cell_spec(ifelse(round(`Pr(>F)`,4)<.001,"< .001",round(`Pr(>F)`,4)),
#                         color = ifelse(is.nan(`Pr(>F)`),"blue", ifelse(`Pr(>F)` < .05, "green", "red")),
#                         bold = ifelse(is.nan(`Pr(>F)`),F, ifelse(`Pr(>F)` < .05, T,F)),
#                         align = "center")
#   ) %>%
#   dplyr::select(`num Df`, `den Df`, MSE, `F`, ges, `Pr(>F)`) %>%
#   kable(escape = F, digits = c(3,3,3,3,3,3,3,3,3), caption = c("ANOVA | dv = RT in ms")) %>%
#   kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 6))


# RT
dat2plot_all %>%
  nest() %>%
  mutate(
    # stats = map(data, ~ broom::tidy(t.test(RT_Mean ~ eventtype2, data = ., paired = TRUE))),
    stats = map(data, ~ rstatix::t_test(RT_Mean ~ eventtype2, data = ., paired = TRUE, detailed = TRUE)),
    t_test_bf = map(data, ~ extractBF(ttestBF(formula = RT_Mean ~ eventtype2, data = ., 
                                              iterations = num.iter, rscale = sqrt(2)/2))),
    meanval = map(data, ~ .x %>%
                    group_by(eventtype2) %>%
                    summarise(
                      mean_RT_Mean = mean(RT_Mean, na.rm = TRUE),
                      std_RT_Mean = sd(RT_Mean, na.rm = TRUE)
                    ) %>%
                    pivot_wider(names_from = eventtype2, 
                                values_from = c(mean_RT_Mean, std_RT_Mean), 
                                names_sep = "_")
    ),
    Dz = map(data, ~ rstatix::cohens_d(RT_Mean ~ eventtype2, data = ., paired = TRUE))
  ) %>%
  # unnest(cols = c(stats, cohens_d, t_test_bf, meanval))%>%
  unnest(cols = c(Dz))%>%
  dplyr::select(stats,t_test_bf, meanval, effsize) %>%
  unnest(cols = c(stats, t_test_bf, meanval)) %>%
  mutate(BF10 = bf, BF01 = 1/bf, 
         M_RT_noprim = mean_RT_Mean_target_nonprimed,
         SD_RT_noprim = std_RT_Mean_target_nonprimed,
         M_RT_prim = mean_RT_Mean_target_primed,
         SD_RT_prim = std_RT_Mean_target_primed,
         diff=estimate,
         Dz = effsize, t.value=statistic)%>%
  dplyr::select(M_RT_noprim, SD_RT_noprim, M_RT_prim, SD_RT_prim, diff, t.value, df, p, Dz, BF10, BF01)%>%
  # mutate(p = p.adjust(p, method="holm")) %>% # none, holm
  # mutate(p = p.adjust(p, method="none")) %>% # none, holm
  ungroup()%>%
  # mutate(
  #   p = cell_spec(p, color = ifelse(is.nan(p),"blue", ifelse(p < .05, "green", "red")))
  # ) %>%
  mutate(
    p = cell_spec(ifelse(round(p,4)<.001,"< .001",round(p,4)),
                  color = ifelse(is.nan(p),"blue", ifelse(p < .05, "green", "red")),
                  bold = ifelse(is.nan(p),F, ifelse(p < .05, T,F)),
                  align = "center")
  ) %>%
  kbl(escape = F, digits = 3, 
        caption = c("RT | modulation by target type (primed vs nonprimed) | t-tests of difference")) %>%
  kable_styling("striped", full_width = T,bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8))
```

**Hitrate** ~ **priming**

<br> 

```{r lme_HR_lum, fig.show = "hide", warning = FALSE, fig.height=4, fig.width=10}
# Hitrate
dat2plot_all %>%
  nest() %>%
  mutate(
    # stats = map(data, ~ broom::tidy(t.test(Hitrate ~ eventtype2, data = ., paired = FALSE))),
    stats = map(data, ~ rstatix::t_test(Hitrate ~ eventtype2, data = ., paired = TRUE, detailed = TRUE)),
    t_test_bf = map(data, ~ extractBF(ttestBF(formula = Hitrate ~ eventtype2, data = ., iterations = num.iter, rscale = sqrt(2)/2))),
    meanval = map(data, ~ .x %>%
                    group_by(eventtype2) %>%
                    summarise(
                      mean_HR_Mean = mean(Hitrate, na.rm = TRUE),
                      std_HR_Mean = sd(Hitrate, na.rm = TRUE)
                    ) %>%
                    pivot_wider(names_from = eventtype2, 
                                values_from = c(mean_HR_Mean, std_HR_Mean), 
                                names_sep = "_")
    ),
    cohens_d = map(data, ~ rstatix::cohens_d(Hitrate ~ eventtype2, data = ., paired = FALSE))
  ) %>%
  # unnest(cols = c(stats, cohens_d, t_test_bf, meanval))%>%
  unnest(cols = c(cohens_d))%>%
  dplyr::select(stats,t_test_bf, meanval, effsize) %>%
  unnest(cols = c(stats, t_test_bf, meanval)) %>%
  mutate(BF10 = bf, BF01 = 1/bf, 
         M_HR_noprim = mean_HR_Mean_target_nonprimed,
         SD_HR_noprim = std_HR_Mean_target_nonprimed,
         M_HR_prim = mean_HR_Mean_target_primed,
         SD_HR_prim = std_HR_Mean_target_primed,
         diff=estimate,
         Dz = effsize, t.value=statistic)%>%
  dplyr::select(M_HR_noprim, SD_HR_noprim, M_HR_prim, SD_HR_prim, diff, t.value, df, p, Dz, BF10, BF01)%>%
  # mutate(p = p.adjust(p, method="holm")) %>% # none, holm
  # mutate(p = p.adjust(p, method="none")) %>% # none, holm
  ungroup()%>%
  mutate(
    p = cell_spec(ifelse(round(p,4)<.001,"< .001",round(p,4)),
                        color = ifelse(is.nan(p),"blue", ifelse(p < .05, "green", "red")),
                        bold = ifelse(is.nan(p),F, ifelse(p < .05, T,F)),
                        align = "center")
  ) %>%
  kbl(escape = F, digits = 3, 
        caption = c("Hitrate | modulation by target type (primed vs nonprimed) | t-tests of difference")) %>%
  kable_styling("striped", full_width = T,bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8))
```
<br> 

## Time-resolved Analysis

### Hit Rate

<br> 

#### Statistical test for time-resolved analysis of probability of hits/hit rates
*Do probabilities of Hits for primed and non-primed events differ (differently) across time?*

full model: response ~ eventtype2 * eventonsettimes + (1 | participant)
reduced model: response ~ eventtype2 + eventonsettimes + (1 | participant)

Test rationale: do these models differ significantly and does the interaction term thus allow for a better explanation of the data.

<br> 
```{r stat_hitxpresentationtime_distr, warning = FALSE, fig.height=4, fig.width=6}
# Full model with interaction

# Model with smoothers by condition
# with time x cue interaction
model_gam_interaction <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         eventtype2 = as.factor(eventtype2))%>%
  gam(hit ~ eventtype2 + s(eventonsettimes, by = eventtype2, bs = "cs") + participant,
                 data = ., family = binomial)
  #
# coef(model_gam_interaction)
# summary(model_gam_interaction)
#

# no interaction only main effects
model_gam_TwoMain <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         eventtype2 = as.factor(eventtype2))%>%
  gam(hit ~ eventtype2 + s(eventonsettimes, bs = "cs") + participant, 
                 data = ., family = binomial)

# no effect of time only CUE
model_gam_OneMain <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         eventtype2 = as.factor(eventtype2))%>%
  gam(hit ~ eventtype2  + participant, 
                 data = ., family = binomial)

# no effect of time
model_gam_null <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         eventtype2 = as.factor(eventtype2))%>%
  gam(hit ~  participant, 
                 data = ., family = binomial)

plot(model_gam_interaction, pages = 1)
# anova(model_gam_interaction)


# based on 'marginal effects'
# plot_predictions(model_gam_interaction, condition = c('eventonsettimes', 'eventtype2'),
#                  type = 'response') +
#       labs(y = "Predicted probability of a hit",
#            title = "Average smooth effect of eventonsettimes | GAM",
#            subtitle = "Per priming condition")


anova(model_gam_null,model_gam_OneMain,model_gam_TwoMain,model_gam_interaction) %>%
  tidy()%>%
  kbl(escape = F, digits=3,
      caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:"))%>%
  # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>",
  #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = 
                                                                   "hold_position"))


AIC(model_gam_null, model_gam_OneMain, model_gam_TwoMain, model_gam_interaction) %>%
  # tidy()%>%
  kbl(escape = F, digits=3,
      caption = paste0("AIC | comparing full and reduced model | GAM:"))%>%
  # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>",
  #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options =
                                                                   "hold_position"))
# based on 'gratia'
# # Plot difference between smooths (e.g., eventtype2 levels "A" and "B")
# diff_df <- difference_smooths(model_gam_interaction,
#                                smooth = "s(eventonsettimes)",
#                                comparison = c("eventtype2A", "eventtype2B"),
#                                n = 10)  # 200 points across the range

# diff_df <- diff_df %>%
#   mutate(
#     .diff = plogis(.diff),
#     .lower_c = plogis(.lower_ci),
#     .upper_ci = plogis(.upper_ci)
#   )

# ggplot(diff_df, aes(x = eventonsettimes, y = .diff)) +
#   geom_line() +
#   geom_ribbon(aes(ymin = .lower_ci, ymax = .upper_ci), alpha = 0.3) +
#   geom_hline(yintercept = 0, linetype = "dashed") +
#   labs(x = "Event onset time", y = "Difference in log-odds of Hitrate (nonprimed - primed)")


```

<br> 

##### follow up: directly test time windows | Hitrates
*Do probabilities of Hits for primed and non-primed events differ (differently) across time?*

full model: response ~ eventtype2 * eventonsettimes + (1 | participant)

Test rationale: do these models differ significantly and does the interaction term thus allow for a better explanation of the data.

Extract predicted probability of hits for early and late window and contrast them

<br> 

```{r stat_hitxpresentationtime_distr_contrasts, warning = FALSE, fig.height=3.5, fig.width=4.9}
# Full model with interaction
model_gam_interaction <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         eventtype2 = as.factor(eventtype2))%>%
  gam(hit ~ eventtype2 + s(eventonsettimes, by = eventtype2, bs = "cs") + participant,
                 data = ., family = binomial)


# alternative approach
# Define the time windows
all_timepoints = sort(unique(DATA_Trials$eventonsettimes))
early_window <- all_timepoints[
  (all_timepoints >= min(unique(DATA_Trials$eventonsettimes)) + 0) & (all_timepoints <= min(unique(DATA_Trials$eventonsettimes)) + 0.5)
]
late_window <-all_timepoints[
  (all_timepoints >= max(unique(DATA_Trials$eventonsettimes)) -.5) & (all_timepoints <= max(unique(DATA_Trials$eventonsettimes)) + 0)
]

emm_predictions_early <- emmeans(
  model_gam_interaction,
  ~ eventtype2 ,
  at = list(eventonsettimes = early_window)
)

emm_predictions_late <- emmeans(
  model_gam_interaction,
  ~ eventtype2 ,
  at = list(eventonsettimes = late_window)
)

# 1. Add the identifier column to the grid of each object
emm_predictions_early@grid$eventonsettimes <- "early"
emm_predictions_late@grid$eventonsettimes  <- "late"


# Combine the objects using named arguments
emm_predictions <- rbind(early = emm_predictions_early, 
                      late  = emm_predictions_late)

summary(emm_predictions, type = "response") %>%
  as.data.frame()%>%
  # mutate(eventonsettimes=c("early","early","late","late"))%>%
  kbl(escape = F, digits=4,
      caption = paste0("Marginal Mean predicted Probabilities in early and late time window"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))

# Now, set up the contrast to compare early vs. late for each condition.
# This will perform a pairwise comparison for each event type.
# do everything on the response scale, i.e. probability scale
emm_response_scale <- regrid(emm_predictions)

# Now perform the contrasts on the regridded object
contrast_tdiffs <- contrast(
  emm_response_scale,
  method = "pairwise",
  by = "eventtype2",
  adjust= "holm"
)


# View the results on the probability scale
summary(contrast_tdiffs) %>%
  as.data.frame()%>%
  mutate(contrast = c("early vs late"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  mutate(`CohensD1` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = df.residual(model_gam_interaction), method = "identity"))$effect.size
  )%>%
  mutate(`CohensD2` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = sum(model_gam_interaction$edf), method = "identity"))$effect.size
  )%>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of hit rates between early and late time window | Holm corrected"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))



# Now perform the contrasts on the regridded object
contrast_condiffs <- contrast(
  emm_response_scale,
  method = "pairwise",
  by = "eventonsettimes",
  adjust= "holm"
)


# View the results on the probability scale
summary(contrast_condiffs) %>%
  as.data.frame()%>%
  mutate(eventonsettimes = c("early","late"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  mutate(`CohensD1` =
           summary(eff_size(contrast_condiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = df.residual(model_gam_interaction), method = "identity"))$effect.size
  )%>%
  mutate(`CohensD2` =
           summary(eff_size(contrast_condiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = sum(model_gam_interaction$edf), method = "identity"))$effect.size
  )%>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of hit rates between conditions | Holm corrected"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))


contrast(
  emm_response_scale,
  method = "revpairwise",
  by = "eventonsettimes"
) %>% 
  contrast(method = "pairwise", by = NULL) %>%
  summary()%>%
  as.data.frame()%>%
  mutate(contrast = c("early(primed-nonprimed) - late(primed-nonprimed)"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of hit rates diff of diffs "))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))




emm_predictions_plot <- emm_predictions%>%
  regrid()%>%
  as.data.frame() %>%
  mutate(eventonsettimes = case_when(
    eventonsettimes=="early" ~ mean(early_window),
    eventonsettimes=="late" ~ mean(late_window),
  ))




# Generate a data frame of predictions for the eventonsettimes effect
predictions_df <- ggpredict(
  model_gam_interaction,
  terms = c("eventonsettimes","eventtype2")
)


 
plotplot <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         eventtype2 = as.factor(eventtype2))%>%
  ggplot(., aes(x = eventonsettimes, y = as.numeric(hit), color = eventtype2)) +
  # Original data points
  stat_summary(fun = mean, geom = "point", alpha = 0.4, bins=length(unique(DATA_Trials$eventonsettimes))) +
   # Add the GAM predictions as a line and a ribbon
  # geom_ribbon(data = predictions_df, aes(x = x, y = predicted,ymin = conf.low, ymax = conf.high, fill = group),
  #             alpha = 0.4, , colour = NA) +
  geom_ribbon(data = predictions_df, aes(x = x, y = predicted,fill = group, ymin = conf.low, ymax = conf.high, group=group),
              alpha = 0.2, , colour = NA) +
  # geom_ribbon(data = predictions_df, aes(x = x, y = predicted,ymin = conf.low, ymax = conf.high), fill = "steelblue", alpha = 0.2) +
  geom_line(data = predictions_df, aes(x = x, y = predicted, color = group), size = 1)+
    # Labels and theme
  labs(x = "Event Onset Time (s)", y = "Probability of Hits",
       title = "Probability of Hits | GAM model", subtitle = "+ predicted hit rates for windows of interest") +
  geom_point(data = emm_predictions_plot, aes(y = prob, group = eventtype2), size = 2, stroke = 1) +
  geom_errorbar(data = emm_predictions_plot, aes(y = prob,ymin = lower.CL, ymax = upper.CL), width = 0.1, size = 1) +
  scale_fill_manual( values=c("#198a83", "#ff8504")) +
  scale_color_manual( values=c("#198a83", "#ff8504")) +
  # Add vertical lines for the early time window
  annotate("segment", x = early_window[1], xend = rev(early_window)[1], y = 0.3, yend = 0.3, color = "grey60", size = 1.5) +
  annotate("segment", x = late_window[1], xend = rev(late_window)[1], y = 0.3, yend = 0.3, color = "grey60", size = 1.5) +
  # theme_minimal()
  theme_light()


savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", height = 3,width = 4.9,  unit = "in")
savefig %<>%
  fill_panel(plotplot, column = 1, row = 1)
savefig
# save_multi_panel_figure(savefig, "figures/Behavior_ProbHits.pdf")

```

<br> 

### Reaction Times

#### Statistical test for time-resolved analysis of reaction times
*Do reaction times for hits for primed and non-primed events differ (differently) across time?*

full model: RT ~ eventtype2 * eventonsettimes + (1 | participant)

reduced model: RT ~ eventtype2 + eventonsettimes + (1 | participant)

<br> 
```{r stat_RTxpresentationtime_distr,  warning = FALSE, fig.height=4, fig.width=6}
# Model with smoothers by condition
# with time x cue interaction
# Model with smoothers by condition
model_gam_interaction <- DATA_Trials %>%
  filter(response=='hit')%>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(RT ~ eventtype2 + s(eventonsettimes, by = eventtype2, bs = "cs") + participant, 
                 data = .)
# coef(model_gam_interaction)
# summary(model_gam_interaction)
# gam.check(model_gam_interaction)


# no interaction only main effects
model_gam_TwoMain <- DATA_Trials %>%
filter(response=='hit')%>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(RT ~ eventtype2 + s(eventonsettimes, bs = "cs") + participant, 
                 data = .)

# no effect of time only CUE
model_gam_OneMain <- DATA_Trials %>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(RT ~ eventtype2 +  participant, 
                 data = .)

# no effect of time
model_gam_null <- DATA_Trials %>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(RT ~ participant, 
                 data = .)
  
plot(model_gam_interaction, pages = 1)
# anova(model_gam_interaction)


# based on 'marginal effects'
# plot_predictions(model_gam_interaction, condition = c('eventonsettimes', 'eventtype2'),
#                  type = 'response') +
#       labs(y = "Predicted probability of a hit",
#            title = "Average smooth effect of eventonsettimes | GAM",
#            subtitle = "Per priming condition")


anova(model_gam_null,model_gam_OneMain,model_gam_TwoMain,model_gam_interaction) %>%
  tidy()%>%
  kbl(escape = F, digits=3,
      caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:"))%>%
  # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>",
  #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = 
                                                                   "hold_position"))


AIC(model_gam_null, model_gam_OneMain, model_gam_TwoMain, model_gam_interaction) %>%
  # tidy()%>%
  kbl(escape = F, digits=3,
      caption = paste0("AIC | comparing full and reduced model | GAM:"))%>%
  # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>",
  #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options =
                                                                   "hold_position"))
```

<br> 


##### follow up: directly test time windows | RTs
*Do RTs for primed and non-primed events differ (differently) across time?*

full model: RT ~ eventtype2 * eventonsettimes + (1 | participant)

Test rationale: do these models differ significantly and does the interaction term thus allow for a better explanation of the data.

Extract predicted probability of RTs for early and late window and contrast them

<br> 
```{r stat_RTxpresentationtime_distr_contrasts, warning = FALSE, fig.height=3.5, fig.width=4.9}
# Full model with interaction
# Model with smoothers by condition
model_gam_interaction <- DATA_Trials %>%
  filter(response=='hit')%>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(RT ~ eventtype2 + s(eventonsettimes, by = eventtype2, bs = "cs") + participant, 
                 data = .)

# alternative approach
# Define the time windows
all_timepoints = sort(unique(DATA_Trials$eventonsettimes))
early_window <- all_timepoints[
  (all_timepoints >= min(unique(DATA_Trials$eventonsettimes)) + 0) & (all_timepoints <= min(unique(DATA_Trials$eventonsettimes)) + 0.5)
]
late_window <-all_timepoints[
  (all_timepoints >= max(unique(DATA_Trials$eventonsettimes)) -.5) & (all_timepoints <= max(unique(DATA_Trials$eventonsettimes)) + 0)
]

emm_predictions_early <- emmeans(
  model_gam_interaction,
  ~ eventtype2 ,
  at = list(eventonsettimes = early_window)
)

emm_predictions_late <- emmeans(
  model_gam_interaction,
  ~ eventtype2 ,
  at = list(eventonsettimes = late_window)
)

# 1. Add the identifier column to the grid of each object
emm_predictions_early@grid$eventonsettimes <- "early"
emm_predictions_late@grid$eventonsettimes  <- "late"


# Combine the objects using named arguments
emm_predictions <- rbind(early = emm_predictions_early, 
                      late  = emm_predictions_late)

summary(emm_predictions, type = "response") %>%
  as.data.frame()%>%
  # mutate(eventonsettimes=c("early","early","late","late"))%>%
  kbl(escape = F, digits=4,
      caption = paste0("Marginal Mean predicted RTs in early and late time window"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))

# Now, set up the contrast to compare early vs. late for each condition.
# This will perform a pairwise comparison for each event type.
# do everything on the response scale, i.e. probability scale
emm_response_scale <- regrid(emm_predictions)

# Now perform the contrasts on the regridded object
contrast_tdiffs <- contrast(
  emm_response_scale,
  method = "pairwise",
  by = "eventtype2",
  adjust= "holm"
)


# View the results on the probability scale
summary(contrast_tdiffs) %>%
  as.data.frame()%>%
  mutate(contrast = c("early vs late"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  mutate(`CohensD1` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = df.residual(model_gam_interaction), method = "identity"))$effect.size
  )%>%
  mutate(`CohensD2` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = sum(model_gam_interaction$edf), method = "identity"))$effect.size
  )%>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of RTs between early and late time window | Holm corrected"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))


# Now perform the contrasts on the regridded object
contrast_condiffs <- contrast(
  emm_response_scale,
  method = "pairwise",
  by = "eventonsettimes",
  adjust= "holm"
)


# View the results on the probability scale
summary(contrast_condiffs) %>%
  as.data.frame()%>%
  mutate(eventonsettimes = c("early","late"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  mutate(`CohensD1` =
           summary(eff_size(contrast_condiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = df.residual(model_gam_interaction), method = "identity"))$effect.size
  )%>%
  mutate(`CohensD2` =
           summary(eff_size(contrast_condiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = sum(model_gam_interaction$edf), method = "identity"))$effect.size
  )%>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of RTs between conditions | Holm corrected"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))


emm_predictions_plot <- emm_predictions%>%
  regrid()%>%
  as.data.frame() %>%
  mutate(eventonsettimes = case_when(
    eventonsettimes=="early" ~ mean(early_window),
    eventonsettimes=="late" ~ mean(late_window),
  ))



# Generate a data frame of predictions for the eventonsettimes effect
predictions_df <- ggpredict(
  model_gam_interaction,
  terms = c("eventonsettimes","eventtype2")
)

plotplot <- DATA_Trials %>%
  filter(response=='hit')%>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  ggplot(., aes(x = eventonsettimes, y = as.numeric(RT), color = eventtype2)) +
  # Original data points
  stat_summary(fun = mean, geom = "point", alpha = 0.4, bins=length(unique(DATA_Trials$eventonsettimes))) +
  
  # Add the GAM predictions as a line and a ribbon
  # geom_ribbon(data = predictions_df, aes(x = x, y = predicted,ymin = conf.low, ymax = conf.high, fill = group),
  #             alpha = 0.4, , colour = NA) +
  geom_ribbon(data = predictions_df, aes(x = x, y = predicted,fill = group, ymin = conf.low, ymax = conf.high, group=group),
              alpha = 0.2, , colour = NA) +
  # geom_ribbon(data = predictions_df, aes(x = x, y = predicted,ymin = conf.low, ymax = conf.high), fill = "steelblue", alpha = 0.2) +
  geom_line(data = predictions_df, aes(x = x, y = predicted, color = group), size = 1)+
    # Labels and theme
  labs(x = "Event Onset Time (s)", y = "RT",
       title = "RT | GAM model", subtitle = "+ predicted RT for windows of interest") +
  geom_point(data = emm_predictions_plot, aes(y = emmean, group = eventtype2), size = 2,  stroke = 1) +
  geom_errorbar(data = emm_predictions_plot, aes(y = emmean,ymin = lower.CL, ymax = upper.CL), width = 0.1, size = 1) +
  scale_fill_manual( values=c("#198a83", "#ff8504")) +
  scale_color_manual( values=c("#198a83", "#ff8504")) +
  # Add vertical lines for the early time window
  annotate("segment", x = early_window[1], xend = rev(early_window)[1], y = 500, yend = 500, color = "grey60", size = 1.5) +
  annotate("segment", x = late_window[1], xend = rev(late_window)[1], y = 500, yend = 500, color = "grey60", size = 1.5) +
  # theme_minimal()+
  theme_light() +
  guides(fill="none")

savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", height = 3,width = 4.9,  unit = "in")
savefig %<>%
  fill_panel(plotplot, column = 1, row = 1)
savefig
# save_multi_panel_figure(savefig, "figures/Behavior_ProbRT.pdf")

```
<br> 


<br> 

### False Alarm Rates

#### Display of smoothed False Alarm Rates across time
What are the distributions of false alarm rates across the trial?

<br> 

```{r disp_FA_X_TIME, fig.height=4, fig.width=7, warning = FALSE}
DATA_FA <- DataIn1%>%
  filter(eventtype=="distractor")%>%
  mutate(FA = ifelse(response == "FA_proper", 1, 0))

ggplot(DATA_FA,aes(x = eventonsettimes, y = FA)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "loess", se = TRUE, span = 0.3) +    # kernel smoothed average
  labs(
    title = "Smoothed FA probabilities",
    x = "eventonsettimes (s)",
    y = "FA probability"
  ) +
  theme_minimal()

ggplot(DATA_FA, aes(x = eventonsettimes, y = FA)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), se = TRUE) +
  labs(x = "Event Onset Time", y = "FA probability", 
       title = "GAM Smoothed Probability of FAs") +
  theme_minimal()



```
<br> 

#### Statistical test for time-resolved analysis of FArate
*Does the probability of FAs differ across time?*

full model: FA ~  eventonsettimes + (1 | participant)
reduced model: FA ~  1 + (1 | participant)

Test rationale: do these models differ significantly and does the interaction term thus allow for a better explanation of the data.

<br> 
```{r stat_FA_X_TIME_test, warning = FALSE, fig.height=4, fig.width=6}


# Full model with factor
GLMHit_full <- DataIn1%>%
  filter(eventtype=="distractor")%>%
  mutate(response = as.factor(response)) %>%
  glmer(response ~ eventonsettimes + (1 | participant),data=., family = binomial)

# Reduced model intercept model
GLMHit_reduced <- DataIn1%>%
  filter(eventtype=="distractor")%>%
  mutate(response = as.factor(response)) %>%
  glmer(response ~ 1 + (1 | participant),data=., family = binomial)

# Likelihood ratio test
anova( GLMHit_reduced, GLMHit_full, test = "Chisq") %>%
  as.data.frame() %>%
  kable(escape = F, digits=3,
        caption = paste0("Likelihood ratio test via anova | comparing full and reduced model:<br>", 
                    deparse(formula(GLMHit_full))))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))

# Display the summary of the full model + Wald test
GLMHit_full %>%
  tidy() %>%
  kable(escape = F, digits=3,
        caption = paste0("Likelihood ratio test via anova | comparing full and reduced model:<br>", 
                    deparse(formula(GLMHit_full)))) %>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8))


# Model with smoothers by condition
DATA_FA <- DataIn1%>%
  filter(eventtype=="distractor")%>%
  mutate(FA = ifelse(response == "FA_proper", 1, 0))



model_gam_factor <- DATA_FA%>%
  gam(FA ~ s(eventonsettimes, bs = "cs") + participant, 
                 data = ., family = binomial)

model_gam_null <-DATA_FA%>%
  gam(FA ~ participant, 
                 data = ., family = binomial)

plot(model_gam_factor, pages = 1)
# anova(model_gam_interaction)


# plot_predictions(model_gam_factor, condition = 'eventonsettimes',
#                  type = 'response') +
#       labs(y = "Predicted probability of a FA",
#            title = "Average smooth effect of eventonsettimes | GAM")


# Likelihood ratio test
anova( model_gam_null, model_gam_factor, test = "Chisq") %>%
  tidy()%>%
  # as.data.frame() %>%
  # mutate(model = c('reduced model','full model')) %>%
  # dplyr::select(model, `Resid. Df`:`Pr(>Chi)`)%>%
  kbl(escape = F, digits=3,
      caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))



ggplot(DATA_FA, aes(x = eventonsettimes, y = FA)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"),family = binomial,  se = TRUE) +
  labs(x = "Event Onset Time", y = "FA probability", 
       title = "GAM Smoothed Probability of FAs") +
  theme_minimal()


# Generate a data frame of predictions for the eventonsettimes effect
predictions_df <- ggpredict(
  model_gam_factor,
  terms = "eventonsettimes"
)



```

<br> 

##### follow up: directly test time windows | FAs
*Do FAs for primed and non-primed events differ (differently) across time?*

full model: FA ~ eventtype2 * eventonsettimes + (1 | participant)

Test rationale: do these models differ significantly and does the interaction term thus allow for a better explanation of the data.

Extract predicted probability of FAs for early and late window and contrast them

<br> 
```{r stat_FAxpresentationtime_distr_contrasts, warning = FALSE, fig.height=4, fig.width=7}
# Full model with interaction
# Model with smoothers by condition
model_gam_interaction <- DataIn1%>%
  filter(eventtype=="distractor")%>%
  mutate(FA = ifelse(response == "FA_proper", 1, 0))%>%
  gam(FA ~ s(eventonsettimes, bs = "cs") + participant, 
                 data = ., family = binomial)



# # based on gam interactions
# # Define the time windows
# early_window <- c(0, 0.5) + min(DataIn1$eventonsettimes)
# late_window <- c(-0.5, 0) + max(DataIn1$eventonsettimes)
# 
# # Calculate the average predicted hit probability for each time window
# # and for each event type (attended vs. unattended).
# # We'll use the 'at' argument to specify the time points of interest.
# emm_predictions <- emmeans(
#   model_gam_interaction,
#   ~ eventonsettimes,
#   at = list(eventonsettimes = c(mean(early_window), mean(late_window)))
# )


# alternative approach
# Define the time windows
all_timepoints = sort(unique(DATA_Trials$eventonsettimes))
early_window <- all_timepoints[
  (all_timepoints >= min(unique(DATA_Trials$eventonsettimes)) + 0) & (all_timepoints <= min(unique(DATA_Trials$eventonsettimes)) + 0.5)
]
late_window <-all_timepoints[
  (all_timepoints >= max(unique(DATA_Trials$eventonsettimes)) -.5) & (all_timepoints <= max(unique(DATA_Trials$eventonsettimes)) + 0)
]

emm_predictions_early <- emmeans(
  model_gam_interaction,
  ~ 1 ,
  at = list(eventonsettimes = early_window)
)

emm_predictions_late <- emmeans(
  model_gam_interaction,
  ~ 1 ,
  at = list(eventonsettimes = late_window)
)

# 1. Add the identifier column to the grid of each object
emm_predictions_early@grid$eventonsettimes <- "early"
emm_predictions_late@grid$eventonsettimes  <- "late"


# Combine the objects using named arguments
emm_predictions <- rbind(early = emm_predictions_early, 
                      late  = emm_predictions_late)




summary(emm_predictions, type = "response") %>%
  as.data.frame()%>%
  # mutate(eventonsettimes=c("early","late"))%>%
  kbl(escape = F, digits=4,
      caption = paste0("Marginal Mean predicted RTs in early and late time window"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))


# Now, set up the contrast to compare early vs. late for each condition.
# This will perform a pairwise comparison for each event type.
# do everything on the response scale, i.e. probability scale
emm_response_scale <- regrid(emm_predictions)

# Now perform the contrasts on the regridded object
contrast_tdiffs <- contrast(
  emm_response_scale,
  method = "pairwise"
)


# View the results on the probability scale
summary(contrast_tdiffs) %>%
  as.data.frame()%>%
  mutate(contrast = c("early vs late"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  mutate(`CohensD1` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = df.residual(model_gam_interaction), method = "identity"))$effect.size
  )%>%
  mutate(`CohensD2` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = sum(model_gam_interaction$edf), method = "identity"))$effect.size
  )%>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of FAs between early and late time window"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))


# Now perform the contrasts on the regridded object
contrast_condiffs <- contrast(
  emm_response_scale,
  method = "pairwise"
)


emm_predictions_plot <- emm_predictions%>%
  regrid()%>%
  as.data.frame() %>%
  mutate(eventonsettimes = case_when(
    eventonsettimes=="early" ~ mean(early_window),
    eventonsettimes=="late" ~ mean(late_window),
  ))


# Base plot with original data and GAM smoothed line
 DataIn1%>%
  filter(eventtype=="distractor")%>%
  mutate(FA = ifelse(response == "FA_proper", 1, 0))%>%
  ggplot(., aes(x = eventonsettimes, y = as.numeric(FA))) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +
  # stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.02, alpha = 0.4) + # Optional: add standard error bars
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), se = TRUE) +
  labs(x = "Event Onset Time (s)", y = "Probability of Hits",
       title = "Probability of Hits | GAM model", subtitle = "+ predicted hit rates for windows of interest") +
  geom_point(data = emm_predictions_plot, aes(y = prob), size = 3, shape = 23, stroke = 1.5) +
  geom_errorbar(data = emm_predictions_plot, aes(y = prob,ymin = lower.CL, ymax = upper.CL), width = 0.1, size = 1) +
  # Add vertical lines for the early time window
  annotate("segment", x = early_window[1], xend = early_window[2], y = -0.01, yend = -0.01, color = "grey60", size = 1.5) +
  annotate("segment", x = late_window[1], xend = late_window[2], y = -0.01, yend = -0.01, color = "grey60", size = 1.5) +
  theme_minimal()


# Generate a data frame of predictions for the eventonsettimes effect
predictions_df <- ggpredict(
  model_gam_factor,
  terms = "eventonsettimes"
)

ggplot(DATA_FA, aes(x = eventonsettimes, y = FA)) +
  # Original data points
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +
  # Add the GAM predictions as a line and a ribbon
  geom_line(data = predictions_df, aes(x = x, y = predicted), color = "grey60", size = 1) +
  geom_ribbon(data = predictions_df, aes(x = x, y = predicted,ymin = conf.low, ymax = conf.high), fill = "grey60", alpha = 0.2) +
  # Labels and theme
  labs(x = "Event Onset Time (s)", y = "Probability of FAs",
       title = "Probability of FAs | GAM model", subtitle = "+ predicted FA rates for windows of interest") +
  geom_point(data = emm_predictions_plot, aes(y = prob), size = 2,  stroke =1) +
  geom_errorbar(data = emm_predictions_plot, aes(y = prob,ymin = lower.CL, ymax = upper.CL), width = 0.1, size = 1) +
  # Add vertical lines for the early time window
  annotate("segment", x = early_window[1], xend = rev(early_window)[1], y = -0.01, yend = -0.01, color = "grey60", size = 1.5) +
  annotate("segment", x = late_window[1], xend = rev(late_window)[1], y = -0.01, yend = -0.01, color = "grey60", size = 1.5) +
  theme_light()




```

<br> 

### Control Analysis: Are there differences in the distributions of event times for both cueing conditons?

Idea: can we really compare primed and non-primed events? Are their distributions of presentation times comparable?

Rationale: Can the eventtype be predicted by the presentation time? If presentation time was comparable, it shouldn't be a predictor.

eventtype2 ~ eventonsettimes + (1 | participant)

<br> 
```{r stat_hitxpresentationtime_distr_control, fig.show = "hide", warning = FALSE, fig.height=4, fig.width=10}
GLMEventType_full <- DATA_Trials %>%
  mutate(eventtype2 = as.factor(eventtype2)) %>%
  glmer(eventtype2 ~ eventonsettimes + (1 | participant),data=., family = binomial)


# Display the summary of the full model + Wald test
# summary(GLMEventType_full)

# Display the summary of the full model + Wald test
GLMEventType_full %>%
  tidy() %>%
  kable(escape = F, digits=3,
        caption = paste0("Likelihood ratio test via anova | display of full model:<br>", 
                    deparse(formula(GLMEventType_full)))) %>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8))

```

<br> 

### Control Analysis: Is the sequential modulation simply capturing a sequential strategy in shifting attention by the to be learned cue?

Idea: Can the sequential modulation be solely attributed to the functional classification primed vs non-primed? Or could an alternative account state that the sequential modulation is capturing a sequential strategy in shifting attention by the to be learned cue? Does it even make a difference? If it is learning related, maybe the effect is time-on-task dependent and initially higher than later on.


Rationale: Do a time-on-task resolved analysis

eventtype2 ~ eventonsettimes + (1 | participant)

<br> 
```{r stat_control_learning, fig.height=4, fig.width=8, warning = FALSE}
data_bin <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0))%>%
  group_by(participant)%>%
  # mutate(time_bin = cut(trialnumber,
  #                       breaks = seq(0, max(trialnumber), by = 2),
  #                       labels = FALSE)) %>%
  mutate(time_bin = as.factor(ntile(trialnumber, 2))) %>% 
  ungroup()



ggplot(data_bin, aes(x = eventonsettimes, y = hit, color = eventtype2)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "loess", span = 0.3, se = TRUE) +
  facet_grid(.~time_bin)+
  labs(x = "Event Onset Time", y = "Probability of hits", 
       title = "Smoothed Probability of hits | single sub data") +
  theme_minimal()

# ggplot(data_avg, aes(x = eventonsettimes, y = hit_rate, color = eventtype2)) +
#   # geom_line(alpha = 0.4) +
#   stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
#   geom_smooth(method = "loess", se = TRUE, span = 0.3) +
#   labs(x = "Event Onset Time", y = "Hit Rate", 
#        title = "Smoothed Hit Rates Over Time by Condition | averaged across subjects hit rates") +
#   theme_minimal()


ggplot(data_bin, aes(x = eventonsettimes, y = hit, color = eventtype2)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), se = TRUE) +
  facet_grid(.~time_bin)+
  labs(x = "Event Onset Time", y = "Probability of hits", 
       title = "GAM Smoothed Probability of Hits by Condition") +
  theme_minimal()



data_bin %>%
  filter(response=='hit')%>%
  ggplot(aes(x = eventonsettimes, y = RT, color = eventtype2)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "loess", se = TRUE, span = 0.3) +    # kernel smoothed average
  facet_grid(.~time_bin)+
  labs(
    title = "Smoothed Average Reaction Times",
    x = "eventonsettimes (s)",
    y = "Reaction Time (ms)",
    color = "Condition"
  ) +
  theme_minimal()

data_bin %>%
  filter(response=='hit')%>%
  ggplot(aes(x = eventonsettimes, y = RT, color = eventtype2)) +
  stat_summary(fun = mean, geom = "point", alpha = 0.4) +  # mean RTs
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), se = TRUE) +    # kernel smoothed average
  facet_grid(.~time_bin)+
  labs(
    title = "GAM Smoothed Average Reaction Times",
    x = "eventonsettimes (s)",
    y = "Reaction Time (ms)",
    color = "Condition"
  ) +
  theme_minimal()



# hit rates
# Model with smoothers by condition
model_gam_null <- data_bin%>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(hit ~ eventtype2 + s(eventonsettimes, by = eventtype2, bs = "cs") + participant, 
                 data = ., family = binomial)
# coef(model_gam_interaction)



model_gam_timeOntask <- data_bin%>%
  mutate(
    eventtype2 = as.factor(eventtype2),
    time_bin   = as.factor(time_bin),
    eventtype2_timebin = interaction(eventtype2, time_bin, drop = TRUE)
  ) %>%
  gam(hit ~ eventtype2*time_bin + 
        s(eventonsettimes, by = eventtype2_timebin, bs = "cs") + 
        participant, 
                 data = ., family = binomial)

plot(model_gam_timeOntask, pages = 1)
# anova(model_gam_interaction)


# plot_predictions(model_gam_timeOntask, condition = c('eventonsettimes', 'eventtype2_timebin'),
#                  type = 'response') +
#       labs(y = "Predicted probability of a hit",
#            title = "Average smooth effect of eventonsettimes | GAM",
#            subtitle = "Per priming condition")


# Likelihood ratio test
anova( model_gam_null, model_gam_timeOntask, test = "Chisq") %>%
  tidy()%>%
  # as.data.frame() %>%
  # mutate(model = c('reduced model','full model')) %>%
  # dplyr::select(model, `Resid. Df`:`Pr(>Chi)`)%>%
  kbl(escape = F, digits=3,
      caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))




# reaction times
# Model with smoothers by condition
model_gam_null <- data_bin%>%
  filter(response=='hit')%>%
  mutate(eventtype2 = as.factor(eventtype2))%>%
  gam(RT ~ eventtype2 + s(eventonsettimes, by = eventtype2, bs = "cs") + participant, 
                 data = .)
# coef(model_gam_interaction)



model_gam_timeOntask <- data_bin%>%
  filter(response=='hit')%>%
  mutate(
    eventtype2 = as.factor(eventtype2),
    time_bin   = as.factor(time_bin),
    eventtype2_timebin = interaction(eventtype2, time_bin, drop = TRUE)
  ) %>%
  gam(RT ~ eventtype2*time_bin + 
        s(eventonsettimes, by = eventtype2_timebin, bs = "cs") + 
        participant, 
                 data = .)


plot(model_gam_timeOntask, pages = 1)
# anova(model_gam_interaction)


# plot_predictions(model_gam_timeOntask, condition = c('eventonsettimes', 'eventtype2_timebin'),
#                  type = 'response') +
#       labs(y = "RT",
#            title = "Average smooth effect of eventonsettimes | GAM",
#            subtitle = "Per priming condition")


# Likelihood ratio test
anova( model_gam_null, model_gam_timeOntask, test = "Chisq") %>%
  tidy()%>%
  # as.data.frame() %>%
  # mutate(model = c('reduced model','full model')) %>%
  # dplyr::select(model, `Resid. Df`:`Pr(>Chi)`)%>%
  kbl(escape = F, digits=3,
      caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))

```

<br> 

## Time-resolved Analysis

### SSVEPs

<br> 

#### Statistical test for time-resolved analysis of SSVEP modulations
*Do SSVEP modulations for primed and non-primed events differ (differently) across time?*

full model: response ~ eventtype2 * eventonsettimes + (1 | participant)
reduced model: response ~ eventtype2 + eventonsettimes + (1 | participant)

Test rationale: do these models differ significantly and does the interaction term thus allow for a better explanation of the data.

<br> 
```{r stat_hitxpresentationtime_distr, warning = FALSE, fig.height=4, fig.width=6}
# Full model with interaction

# Model with smoothers by condition
# with time x cue interaction
model_gam_interaction <- DataIn3 %>%
  gam(SSVEP_mod ~ condition + s(time, by = condition ,bs = "cs") + participant, 
                 data = .)

  #
# coef(model_gam_interaction)
# summary(model_gam_interaction)

# no interaction only main effects
model_gam_TwoMain <- DataIn3 %>%
  gam(SSVEP_mod ~ condition + s(time, bs = "cs") + participant, 
                 data = .)

# no effect of time only CUE
model_gam_OneMain <- DataIn3 %>%
  gam(SSVEP_mod ~ condition  + participant, 
                 data = .)

# no effect of time
model_gam_null <- DataIn3 %>%
  gam(SSVEP_mod ~  participant, 
                 data = .)

plot(model_gam_interaction, pages = 1)
# anova(model_gam_interaction)


# based on 'marginal effects'
# plot_predictions(model_gam_interaction, condition = c('eventonsettimes', 'eventtype2'),
#                  type = 'response') +
#       labs(y = "Predicted probability of a hit",
#            title = "Average smooth effect of eventonsettimes | GAM",
#            subtitle = "Per priming condition")


anova(model_gam_null,model_gam_OneMain,model_gam_TwoMain,model_gam_interaction) %>%
  tidy()%>%
  kbl(escape = F, digits=3,
      caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:"))%>%
  # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>",
  #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = 
                                                                   "hold_position"))


AIC(model_gam_null, model_gam_OneMain, model_gam_TwoMain, model_gam_interaction) %>%
  # tidy()%>%
  kbl(escape = F, digits=3,
      caption = paste0("AIC | comparing full and reduced model | GAM:"))%>%
  # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>",
  #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options =
                                                                   "hold_position"))
# based on 'gratia'
# # Plot difference between smooths (e.g., eventtype2 levels "A" and "B")
# diff_df <- difference_smooths(model_gam_interaction,
#                                smooth = "s(eventonsettimes)",
#                                comparison = c("eventtype2A", "eventtype2B"),
#                                n = 10)  # 200 points across the range

# diff_df <- diff_df %>%
#   mutate(
#     .diff = plogis(.diff),
#     .lower_c = plogis(.lower_ci),
#     .upper_ci = plogis(.upper_ci)
#   )

# ggplot(diff_df, aes(x = eventonsettimes, y = .diff)) +
#   geom_line() +
#   geom_ribbon(aes(ymin = .lower_ci, ymax = .upper_ci), alpha = 0.3) +
#   geom_hline(yintercept = 0, linetype = "dashed") +
#   labs(x = "Event onset time", y = "Difference in log-odds of Hitrate (nonprimed - primed)")


```

<br> 

##### follow up: directly test time windows | Hitrates
*Do probabilities of Hits for primed and non-primed events differ (differently) across time?*

full model: response ~ eventtype2 * eventonsettimes + (1 | participant)

Test rationale: do these models differ significantly and does the interaction term thus allow for a better explanation of the data.

Extract predicted probability of hits for early and late window and contrast them

<br> 

```{r stat_hitxpresentationtime_distr_contrasts, warning = FALSE, fig.height=3.5, fig.width=4.9}
# Full model with interaction
model_gam_interaction <- DataIn3 %>%
  gam(SSVEP_mod ~ condition + s(time, by = condition ,bs = "cs") + participant, 
                 data = .)

# alternative approach
# Define the time windows
all_timepoints = sort(unique(DATA_Trials$eventonsettimes))
early_window <- all_timepoints[
  (all_timepoints >= min(unique(DATA_Trials$eventonsettimes)) + 0) & (all_timepoints <= min(unique(DATA_Trials$eventonsettimes)) + 0.5)
]
late_window <-all_timepoints[
  (all_timepoints >= max(unique(DATA_Trials$eventonsettimes)) -.5) & (all_timepoints <= max(unique(DATA_Trials$eventonsettimes)) + 0)
]

emm_predictions_early <- emmeans(
  model_gam_interaction,
  ~ condition ,
  at = list(time = early_window)
)

emm_predictions_late <- emmeans(
  model_gam_interaction,
  ~ condition ,
  at = list(time = late_window)
)

# 1. Add the identifier column to the grid of each object
emm_predictions_early@grid$time <- "early"
emm_predictions_late@grid$time  <- "late"


# Combine the objects using named arguments
emm_predictions <- rbind(early = emm_predictions_early, 
                      late  = emm_predictions_late)

summary(emm_predictions, type = "response") %>%
  as.data.frame()%>%
  # mutate(eventonsettimes=c("early","early","late","late"))%>%
  kbl(escape = F, digits=4,
      caption = paste0("Marginal Mean predicted Probabilities in early and late time window"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))

# Now, set up the contrast to compare early vs. late for each condition.
# This will perform a pairwise comparison for each event type.
# do everything on the response scale, i.e. probability scale
emm_response_scale <- regrid(emm_predictions)

# Now perform the contrasts on the regridded object
contrast_tdiffs <- contrast(
  emm_response_scale,
  method = "pairwise",
  by = "condition",
  adjust= "holm"
)


# View the results on the probability scale
summary(contrast_tdiffs) %>%
  as.data.frame()%>%
  mutate(contrast = c("early vs late"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  mutate(`CohensD1` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = df.residual(model_gam_interaction), method = "identity"))$effect.size
  )%>%
  mutate(`CohensD2` =
           summary(eff_size(contrast_tdiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = sum(model_gam_interaction$edf), method = "identity"))$effect.size
  )%>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of hit rates between early and late time window | Holm corrected"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))



# Now perform the contrasts on the regridded object
contrast_condiffs <- contrast(
  emm_response_scale,
  method = "pairwise",
  by = "time",
  adjust= "holm"
)


# View the results on the probability scale
summary(contrast_condiffs) %>%
  as.data.frame()%>%
  # mutate(time = c("early","late"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  mutate(`CohensD1` =
           summary(eff_size(contrast_condiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = df.residual(model_gam_interaction), method = "identity"))$effect.size
  )%>%
  mutate(`CohensD2` =
           summary(eff_size(contrast_condiffs,
                            sigma = sqrt(mean(sigma(model_gam_interaction, type = "response")^2)), # RMS of sigma()
                            edf = sum(model_gam_interaction$edf), method = "identity"))$effect.size
  )%>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of hit rates between conditions | Holm corrected"))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))


contrast(
  emm_response_scale,
  method = "revpairwise",
  by = "time"
) %>% 
  contrast(method = "pairwise", by = "contrast") %>%
  summary()%>%
  as.data.frame()%>%
  # mutate(contrast = c("early(primed-nonprimed) - late(primed-nonprimed)"))%>%
  mutate(
    p.value = cell_spec(ifelse(round(p.value,4)<.001,"< .001",round(p.value,4)), 
                        color = ifelse(is.nan(p.value),"blue", ifelse(p.value < .05, "green", "red")),
                        bold = ifelse(is.nan(p.value),F, ifelse(p.value < .05, T,F)),
                        align = "center")
  ) %>%
  kbl(escape = F, digits=4,
      caption = paste0("Contrast of marginal means of hit rates diff of diffs "))%>%
        # caption = paste0("Likelihood ratio test via anova | comparing full and reduced model | GAM:<br>", 
        #             paste(deparse(formula(model_gam_interaction)), collapse = " ")))%>%
  kable_styling("striped", full_width = T, bootstrap_options = c("striped", "hover", "condensed", "responsive",font_size = 8,latex_options = "hold_position"))




emm_predictions_plot <- emm_predictions%>%
  regrid()%>%
  as.data.frame() %>%
  mutate(time = case_when(
    time=="early" ~ mean(early_window),
    time=="late" ~ mean(late_window),
  ))




# Generate a data frame of predictions for the eventonsettimes effect
predictions_df <- ggpredict(
  model_gam_interaction,
  terms = c("time","condition")
)


 
plotplot <- DataIn3 %>%
  ggplot(., aes(x = time, y = SSVEP_mod, color = condition)) +
  # Original data points
  stat_summary(fun = mean, geom = "point", alpha = 0.4, bins=length(unique(DATA_Trials$time))) +
   # Add the GAM predictions as a line and a ribbon
  # geom_ribbon(data = predictions_df, aes(x = x, y = predicted,ymin = conf.low, ymax = conf.high, fill = group),
  #             alpha = 0.4, , colour = NA) +
  geom_ribbon(data = predictions_df, aes(x = x, y = predicted,fill = group, ymin = conf.low, ymax = conf.high, group=group),
              alpha = 0.2, , colour = NA) +
  # geom_ribbon(data = predictions_df, aes(x = x, y = predicted,ymin = conf.low, ymax = conf.high), fill = "steelblue", alpha = 0.2) +
  geom_line(data = predictions_df, aes(x = x, y = predicted, color = group), size = 1)+
    # Labels and theme
  labs(x = "Time (s)", y = "SSVEP modulation",
       title = "SSVEP modulation | GAM model", subtitle = "+ predicted SSVEPs for windows of interest") +
  geom_point(data = emm_predictions_plot, aes(y = emmean, group = condition), size = 2, stroke = 1) +
  geom_errorbar(data = emm_predictions_plot, aes(y = emmean,ymin = lower.CL, ymax = upper.CL), width = 0.1, size = 1) +
  scale_fill_manual(values=c("#198A83", "#293C4A", "#F1831A"))+
  scale_color_manual(values=c("#198A83", "#293C4A", "#F1831A")) +
  # Add vertical lines for the early time window
  annotate("segment", x = early_window[1], xend = rev(early_window)[1], y = -15, yend = -15, color = "grey60", size = 1.5) +
  annotate("segment", x = late_window[1], xend = rev(late_window)[1], y = -15, yend = -15, color = "grey60", size = 1.5) +
  # theme_minimal()
  theme_light()


savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", height = 3,width = 4.9,  unit = "in")
savefig %<>%
  fill_panel(plotplot, column = 1, row = 1)
savefig
# save_multi_panel_figure(savefig, "figures/Behavior_ProbHits.pdf")

```

## Time-resolved Analysis linking SSVEPs and behavior

Test rationale:
Use GAM-model predicted data and do a correlation across time for differences smooths

###smoothed differences for hit rate

#### formula
first set up  formula for the smoothed fits later needed for the cross correlation

```{r crosscorrelation_formula, warning = FALSE, fig.height=4, fig.width=6}
get_smooth_vector <- function(df, datainput, time_grid) {
  # 1. Ensure eventtype is a standard factor (NOT ordered)
  df$eventtype_ord <- factor(df$selection, levels = c("stimulus-cued", "top-down-cued"))
  
  # 2. Fit Model (Using 'bam' for speed if it's neural data)
  if(datainput=='HR') {
    # Behavioral HR GAM (Binomial)
    m <- gam(hit ~ eventtype_ord +  
             s(time, by = eventtype_ord, bs = "cs"), 
             data = df, family = binomial(), method = "REML")
    time_var <- "time"
  } else if(datainput=='RT') {
    # Behavioral RT GAM (Gaussian )
    m <- gam(RT ~ eventtype_ord +  
             s(time, by = eventtype_ord, bs = "cs"), 
             data = df, family = gaussian(), method = "REML")
    time_var <- "time"
  } else if(datainput=='SSVEP') {
    # Neural GAM (Gaussian)
    m <- gam(SSVEP_mod ~ eventtype_ord +
             s(time, by = eventtype_ord, bs = "cs"), 
             data = df, family = gaussian(), method = "REML")
    time_var <- "time"
  }
  
  # predict values for the defined time points
  predictions_df <- ggpredict(
    m, 
    terms= list(eventtype_ord = c("stimulus-cued", "top-down-cued"), time=common_times), 
    back_transform = TRUE) %>%
    as.data.frame()%>%
    select(eventtype_ord = x, time = group, predicted) %>%
    pivot_wider(
      names_from = eventtype_ord,
      values_from = predicted
    ) %>%
    mutate(
      diff = `stimulus-cued` - `top-down-cued`
    )
    
  
 
  # return(as.numeric(diff_smooth))
  # return(as.numeric(predictions_df$diff))
  return(predictions_df)
}

```

#### display the cross correlation results and the shifted data

<br> 
```{r correlation_difference_smooths_display, warning = FALSE, fig.height=3, fig.width=6}
#define behavioral and neural data
beh_data <- DATA_Trials %>%
  mutate(hit = ifelse(response == "hit", 1, 0),
         selection = as.factor(eventtype2)) %>%
  mutate(selection = case_when(
    selection=="target_primed" ~ "stimulus-cued",
    selection=="target_nonprimed" ~ "top-down-cued"),
    time = eventonsettimes)%>%
  dplyr::select(participant, selection, time, hit)

neur_data <-  DataIn3 %>%
  mutate(selection = case_when(
    condition=="primed" ~ "stimulus-cued",
    condition=="nonprimed" ~ "top-down-cued"))%>%
  filter(selection=="stimulus-cued"|selection=="top-down-cued")%>%
  dplyr::select(participant, selection, time, SSVEP_mod)

# set up time windows for analysis
# Ensure the time range covers both datasets
common_times <- seq(min(DATA_Trials$eventonsettimes),
                    max(DATA_Trials$eventonsettimes),
                    length.out = 100)


#  extract data
beh_pred <- get_smooth_vector(beh_data, datainput = 'HR', time_grid = common_times)
neu_pred <- get_smooth_vector(neur_data, datainput = 'SSVEP', time_grid = common_times)


# Z-score both for visualization
plot_data <- data.frame(
  time = common_times,
  Behavior = scale(beh_pred$diff),
  SSVEP = scale(neu_pred$diff)
)

ggplot(plot_data, aes(x = time)) +
  geom_line(aes(y = Behavior, color = "Behavioral Diff (Hit Rate)"), size = 1.2) +
  geom_line(aes(y = SSVEP, color = "Neural Diff (SSVEP)"), size = 1.2, linetype = "dashed") +
  theme_minimal() +
  labs(title = "Matching Dynamics: Difference Smooths",
       subtitle = "Z-scored differences (Stimulus-Cued minus Top-Down)",
       y = "Standardized Difference",
       x = "Time (s)",
       color = "Modality") +
  scale_color_manual(values = c("#E41A1C", "#377EB8"))

### cross correlation
# Assuming you have the 'plot_data' from the previous step:
# time: the time grid
# Behavior: Z-scored difference smooth (hit rate)
# SSVEP: Z-scored difference smooth (neural)

# Extract the vectors
beh_vec <- plot_data$Behavior
neu_vec <- plot_data$SSVEP

# Calculate the sampling interval (needed to convert 'lags' to 'milliseconds')
dt <- plot_data$time[2] - plot_data$time[1]

# Run the cross-correlation
# lag.max limits how far left/right we look (e.g., 50 bins)
ccf_result <- ccf(neu_vec, beh_vec, lag.max = 50, plot = FALSE)

# Convert results to a data frame for ggplot
ccf_df <- data.frame(
  Lag_Bins = ccf_result$lag,
  Lag_s = ccf_result$lag * dt,
  Correlation = ccf_result$acf
)

# Find the lag with the maximum correlation
max_lag_s <- ccf_df$Lag_s[which.max(ccf_df$Correlation)]
max_cor <- max(ccf_df$Correlation)

cat("The strongest match occurs when the Brain leads Behavior by", abs(max_lag_s), "s\n")

ggplot(ccf_df, aes(x = Lag_s, y = Correlation)) +
  geom_segment(aes(xend = Lag_s, yend = 0), color = "steelblue") +
  geom_point(color = "darkblue") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") + # Zero lag line
  geom_vline(xintercept = max_lag_s, color = "darkgreen", size = 1) + # Peak lag
  theme_minimal() +
  labs(title = "Cross-Correlation: SSVEP vs. Behavior",
       subtitle = paste("Peak correlation of", round(max_cor, 2), "at", max_lag_s, "s lag"),
       x = "Lag (s) [Negative = Brain leads Behavior]",
       y = "Cross-Correlation Coefficient")



# do cross correlation and plot
# 1. Determine the shift in terms of bins/indices
# If max_lag_s is -150 and dt is 10, shift is -15 bins
shift_bins <- round(max_lag_s / dt)

# 2. Create the shifted data
# # We shift Behavior "backwards" to align with the earlier Neural signal
# plot_data_shifted <- plot_data %>%
#   mutate(
#     Behavior_Shifted = forecast::goldfeld_quandt(Behavior, -shift_bins) # simple shift
#   )

# Manual shift using base R (safer if you don't want extra packages)
shift_vec <- function(x, n) {
  if (n == 0) return(x)
  if (n > 0) return(c(rep(NA, n), head(x, -n))) # Shift right
  return(c(tail(x, n), rep(NA, abs(n))))        # Shift left
}

plot_data$Behavior_Shifted <- shift_vec(plot_data$Behavior, shift_bins)

# 3. Plot the alignment
plotplot<-ggplot(plot_data, aes(x = time)) +
  geom_line(aes(y = SSVEP, color = "Neural Difference"), size = 1.2) +
  geom_line(aes(y = Behavior, color = "Behavioral Difference (Original)"), size = 1.2) +
  geom_line(aes(y = Behavior_Shifted, color = "Behavioral Difference (Lag-Corrected)"), size = 1.2) +
  theme_light() +
  labs(title = "Lag-Corrected Alignment",
       subtitle = paste("Behavior shifted by", max_lag_s, "ms to match Neural dynamics"),
       y = "Z-scored Difference", x = "Time (s)") +
  scale_color_manual(values = c("orange","#E41A1C", "#377EB8"))

savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", height = 3,width = 6,  unit = "in")
savefig %<>%
  fill_panel(plotplot, column = 1, row = 1)
savefig

# save_multi_panel_figure(savefig, "figures/Crosscorr_Curves_HRSSVEP.pdf")

```

#### calculate cross correlation with permutation statistics

```{r calc_crosscorrelation_difference_smooths_permutation, warning = FALSE, fig.height=4, fig.width=6}

# define data

run_permutation <- FALSE  # Set to TRUE to run, FALSE to load from disk
results_filename <- "GAM_Permutation_Results.RData"
n_perm <- 5000 #500 

if (run_permutation) {
  # --- 2. RUN THE LOOP ---
  cat("Running permutations... this may take a while.\n")
  
  n_perm <- 5000 #500 
  null_ccfs <- matrix(NA, nrow = n_perm, ncol = 101) # 101 is default for lag.max=50
  s_beh_perm  <- matrix(NA, nrow = n_perm, ncol = length(common_times)) # 101 is default for lag.max=50
  s_neu_perm  <- s_beh_perm
  s_beh_perm_z <- s_beh_perm
  s_neu_perm_z <- s_beh_perm
  
  # df <- beh_data
  
  
  # # troubleshooting
  # test_diff <- get_diff_smooth_vector(beh_data, is_behavioral = TRUE, time_grid = common_times)
  # test_diff <- get_diff_smooth_vector(neur_data, is_behavioral = FALSE, time_grid = common_times)
  
  for(i in 1:n_perm) {
    # Shuffle labels
    shuff_beh <- beh_data %>% 
      mutate(selection = sample(selection))
    shuff_neu <- neur_data  %>% 
      mutate(selection = sample(selection))
    
    # Extract difference smooths for the shuffled data
    s_beh <- get_smooth_vector(shuff_beh, datainput = 'HR', common_times)
    s_neu <- get_smooth_vector(shuff_neu, datainput = 'SSVEP', common_times)
    
    # Standardize (Z-score) before CCF
    s_beh_z <- as.numeric(scale(s_beh$diff))
    s_neu_z <- as.numeric(scale(s_neu$diff))
    
    # Run CCF
    res <- ccf(s_neu_z, s_beh_z, lag.max = 50, plot = FALSE)
    null_ccfs[i, ] <- res$acf
    s_beh_perm[i, ] <- s_beh$diff
    s_neu_perm[i, ] <- s_neu$diff
    s_beh_perm_z[i, ] <- s_beh_z
    s_neu_perm_z[i, ] <- s_neu_z
    
    if(i %% 50 == 0) cat("Completed permutation", i, "\n")
  }
  
  # --- 3. SAVE TO DISK ---
  save(null_ccfs, s_beh_perm, s_neu_perm, s_beh_perm_z, s_neu_perm_z, 
       file = results_filename)
  cat("Permutations complete. Results saved to:", results_filename, "\n")
} else {
  # --- 4. LOAD PREVIOUSLY SAVED DATA ---
  if (file.exists(results_filename)) {
    load(results_filename)
    cat("Loaded permutation results from:", results_filename, "\n")
  } else {
    stop("Switch is set to 'FALSE' but the results file was not found! Please set switch to 'TRUE' to run permutations once.")
  }
}
  # Calculate 95% Confidence Ribbon
ccf_df$lower_CI <- apply(null_ccfs, 2, quantile, probs = 0.025)
ccf_df$upper_CI <- apply(null_ccfs, 2, quantile, probs = 0.975)


```

#### plot cross correlation with statistics

```{r plot_crosscorrelation_difference_smooths_permutation, warning = FALSE, fig.height=3, fig.width=4}

# Calculate Montecarlo p-values
obs_acf <- as.numeric(ccf_result$acf) # Length is e.g., 101

# 2. Calculate the p-value for each lag
# We compare the absolute values (two-tailed test)
# 'abs(null_ccfs)' is [n_perm x 101]
# 'abs(obs_acf)' is [101]<
# R will recycle obs_acf across the rows of null_ccfs automatically
greater_than_obs <- sweep(abs(null_ccfs), 2, abs(obs_acf), ">=")

# 3. Monte Carlo p-value formula: (Count + 1) / (N + 1)
# colSums gives us one count per lag
p_values <- (colSums(greater_than_obs) + 1) / (n_perm + 1)

# 4. Add to your results data frame
ccf_df$p_val <- p_values


# --- 1. Prepare Permutation Data ---
# Function to convert matrix to long format
matrix_to_long <- function(mat, modality_name, time_vec) {
  as.data.frame(mat) %>%
    mutate(PermID = row_number()) %>%
    pivot_longer(cols = -PermID, names_to = "TimeBin", values_to = "Value") %>%
    mutate(
      # Convert TimeBin (e.g., "V1") to actual time values
      Time = time_vec[as.numeric(gsub("V", "", TimeBin))],
      Modality = modality_name,
      Type = "Permuted"
    )
}

perm_beh_long <- matrix_to_long(s_beh_perm, "Behavior (Hit Rate)", common_times)
perm_neu_long <- matrix_to_long(s_neu_perm, "Neural (SSVEP)", common_times)
perm_beh_long_z <- matrix_to_long(s_beh_perm_z, "Behavior (Hit Rate)", common_times)
perm_neu_long_z <- matrix_to_long(s_neu_perm_z, "Neural (SSVEP)", common_times)

# Combine them
perm_data_all <- bind_rows(perm_beh_long, perm_neu_long)
perm_data_all_z <- bind_rows(perm_beh_long_z, perm_neu_long_z)

# --- 2. Prepare Observed Data ---
obs_diff_beh_z <- as.numeric(scale(get_smooth_vector(beh_data, datainput = 'HR', common_times)$diff))
obs_diff_neu_z <- as.numeric(scale(get_smooth_vector(neur_data, datainput = 'SSVEP', common_times)$diff))
obs_diff_beh <- as.numeric(get_smooth_vector(beh_data, datainput = 'HR', common_times)$diff)
obs_diff_neu <- as.numeric(get_smooth_vector(neur_data, datainput = 'SSVEP', common_times)$diff)

obs_data_all <- data.frame(
  Time = rep(common_times, 2),
  Value_z = c(obs_diff_beh_z, obs_diff_neu_z), # Use the Z-scored versions
  Value = c(obs_diff_beh, obs_diff_neu), # Use the Z-scored versions
  Modality = rep(c("Behavior (Hit Rate)", "Neural (SSVEP)"), each = length(common_times)),
  Type = "Observed"
)

ggplot() +
  # Plot the Permuted Lines (Spaghetti)
  geom_line(data = perm_data_all, 
            aes(x = Time, y = Value, group = interaction(PermID, Modality)), 
            color = "gray80", alpha = 0.2, size = 0.5) +
  
  # Plot the Observed Line (Bold)
  geom_line(data = obs_data_all, 
            aes(x = Time, y = Value, color = Modality), 
            size = 1.2) +
  
  # Facet by Modality to see them side-by-side
  facet_wrap(~Modality, scales = "free_y") +
  
  # Styling
  theme_minimal() +
  scale_color_manual(values = c("Behavior (Hit Rate)" = "#d6604d", "Neural (SSVEP)" = "#4393c3")) +
  labs(
    title = "Observed Dynamics vs. Null Distribution",
    subtitle = "Gray lines represent 5000 permutations with shuffled condition labels",
    x = "Time (s)",
    y = "Difference Smooth in measurement units",
    color = "Modality"
  ) +
  theme(legend.position = "none") # Facet labels already identify the plots

ggplot() +
  # Plot the Permuted Lines (Spaghetti)
  geom_line(data = perm_data_all_z, 
            aes(x = Time, y = Value, group = interaction(PermID, Modality)), 
            color = "gray80", alpha = 0.2, size = 0.5) +
  
  # Plot the Observed Line (Bold)
  geom_line(data = obs_data_all, 
            aes(x = Time, y = Value_z, color = Modality), 
            size = 1.2) +
  
  # Facet by Modality to see them side-by-side
  facet_wrap(~Modality, scales = "free_y") +
  theme_minimal() +
  scale_color_manual(values = c("Behavior (Hit Rate)" = "#d6604d", "Neural (SSVEP)" = "#4393c3")) +
  labs(
    title = "Observed Dynamics vs. Null Distribution",
    subtitle = "Gray lines represent 5000 permutations with shuffled condition labels",
    x = "Time (s)",
    y = "z scored Difference Smooth",
    color = "Modality"
  ) +
  theme(legend.position = "none") # Facet labels already identify the plots



plotplot<-ggplot(ccf_df, aes(x = Lag_s, y = Correlation)) +
  # The Null Distribution Ribbon (from permutations)
  geom_ribbon(aes(ymin = lower_CI, ymax = upper_CI), fill = "grey80", alpha = 0.5) +
  # Zero lines for reference
  geom_hline(yintercept = 0, color = "black", size = 0.3) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  # The observed cross-correlation
  geom_line(color = "#2166ac", size = 1) +
  # Highlight the peak
  # Highlight significant points where p < 0.05
  geom_point(data = subset(ccf_df, p_val < 0.05),
             aes(x = Lag_s, y = Correlation),
             color = "#2166ac", size = 2) +
  geom_segment(aes(x = max_lag_s,
                 xend = max_lag_s,
                 y = 0,
                 yend = max_cor),
             color = "red",
             linewidth = 1) +
  geom_point(aes(x = max_lag_s, y = max_cor), color = "red", size = 3) +
  coord_cartesian(ylim = c(-1, 1)) +
  theme_light() +
  labs(
    title = "Cross-Correlation: SSVEP vs. Behavior",
    subtitle = c(subtitle = paste("Peak correlation of", round(max_cor, 3), "at", max_lag_s, "s lag with p = ", ccf_df$p_val[which.max(ccf_df$Correlation)],
                                  "\nGrey ribbon = 95% null distribution (permutation-based)")),
    x = "Lag (s) [Shift of Behavior relativ to SSVEP]",
    y = "Cross-Correlation (r)"
  )

savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", height = 3,width = 4,  unit = "in")
savefig %<>%
  fill_panel(plotplot, column = 1, row = 1)
savefig

# save_multi_panel_figure(savefig, "figures/Crosscorr_Histogram_HRSSVEP.pdf")
```

###smoothed differences for reaction times

#### display the cross correlation results and the shifted data

<br> 
```{r correlation_difference_smooths_display_RT, warning = FALSE, fig.height=3, fig.width=6}
#define behavioral and neural data
beh_data <- DATA_Trials %>%
  filter(response=='hit')%>%
  mutate(selection = case_when(
    eventtype2=="target_primed" ~ "stimulus-cued",
    eventtype2=="target_nonprimed" ~ "top-down-cued"),
    time = eventonsettimes)%>%
  dplyr::select(participant, selection, time, RT)

neur_data <-  DataIn3 %>%
  mutate(selection = case_when(
    condition=="primed" ~ "stimulus-cued",
    condition=="nonprimed" ~ "top-down-cued"))%>%
  filter(selection=="stimulus-cued"|selection=="top-down-cued")%>%
  dplyr::select(participant, selection, time, SSVEP_mod)

# set up time windows for analysis
# Ensure the time range covers both datasets
common_times <- seq(min(DATA_Trials$eventonsettimes),
                    max(DATA_Trials$eventonsettimes),
                    length.out = 100)
# common_times <- all_timepoints # use time points as above


#  extract data
beh_pred <- get_smooth_vector(beh_data, datainput = 'RT', time_grid = common_times)
neu_pred <- get_smooth_vector(neur_data, datainput = 'SSVEP', time_grid = common_times)


# Z-score both for visualization
plot_data <- data.frame(
  time = common_times,
  Behavior = scale(beh_pred$diff),
  SSVEP = scale(neu_pred$diff)
)

ggplot(plot_data, aes(x = time)) +
  geom_line(aes(y = Behavior, color = "Behavioral Diff (Reaction Time)"), size = 1.2) +
  geom_line(aes(y = SSVEP, color = "Neural Diff (SSVEP)"), size = 1.2, linetype = "dashed") +
  theme_minimal() +
  labs(title = "Matching Dynamics: Difference Smooths",
       subtitle = "Z-scored differences (Stimulus-Cued minus Top-Down)",
       y = "Standardized Difference",
       x = "Time (s)",
       color = "Modality") +
  scale_color_manual(values = c("#E41A1C", "#377EB8"))


### cross correlation
# Assuming you have the 'plot_data' from the previous step:
# time: the time grid
# Behavior: Z-scored difference smooth (hit rate)
# SSVEP: Z-scored difference smooth (neural)

# Extract the vectors
beh_vec <- plot_data$Behavior
neu_vec <- plot_data$SSVEP

# Calculate the sampling interval (needed to convert 'lags' to 'milliseconds')
dt <- plot_data$time[2] - plot_data$time[1]

# Run the cross-correlation
# lag.max limits how far left/right we look (e.g., 50 bins)
ccf_result_RT <- ccf(neu_vec, beh_vec, lag.max = 50, plot = FALSE)

# Convert results to a data frame for ggplot
ccf_df_RT <- data.frame(
  Lag_Bins = ccf_result_RT$lag,
  Lag_s = ccf_result_RT$lag * dt,
  Correlation = ccf_result_RT$acf
)

# Find the lag with the maximum correlation
max_lag_s <- ccf_df_RT$Lag_s[which.max(abs(ccf_df_RT$Correlation))]
max_cor <- ccf_df_RT$Correlation[which.max(abs(ccf_df_RT$Correlation))]

cat("The strongest match occurs when the Brain leads Behavior by", abs(max_lag_s), "s\n")

ggplot(ccf_df_RT, aes(x = Lag_s, y = Correlation)) +
  geom_segment(aes(xend = Lag_s, yend = 0), color = "steelblue") +
  geom_point(color = "darkblue") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") + # Zero lag line
  geom_vline(xintercept = max_lag_s, color = "darkgreen", size = 1) + # Peak lag
  theme_minimal() +
  labs(title = "Cross-Correlation: SSVEP vs. Behavior",
       subtitle = paste("Peak correlation of", round(max_cor, 2), "at", max_lag_s, "s lag"),
       x = "Lag (s) [Negative = Brain leads Behavior]",
       y = "Cross-Correlation Coefficient")



# do cross correlation and plot
# 1. Determine the shift in terms of bins/indices
# If max_lag_s is -150 and dt is 10, shift is -15 bins
shift_bins <- round(max_lag_s / dt)

# 2. Create the shifted data
# # We shift Behavior "backwards" to align with the earlier Neural signal
# plot_data_shifted <- plot_data %>%
#   mutate(
#     Behavior_Shifted = forecast::goldfeld_quandt(Behavior, -shift_bins) # simple shift
#   )

# Manual shift using base R (safer if you don't want extra packages)
shift_vec <- function(x, n) {
  if (n == 0) return(x)
  if (n > 0) return(c(rep(NA, n), head(x, -n))) # Shift right
  return(c(tail(x, n), rep(NA, abs(n))))        # Shift left
}

plot_data$Behavior_Shifted <- shift_vec(plot_data$Behavior, shift_bins)

# 3. Plot the alignment
plotplot<-ggplot(plot_data, aes(x = time)) +
  geom_line(aes(y = SSVEP, color = "Neural Difference"), size = 1.2) +
  geom_line(aes(y = Behavior, color = "Behavioral Difference (Original)"), size = 1.2) +
  geom_line(aes(y = Behavior_Shifted, color = "Behavioral Difference (Lag-Corrected)"), size = 1.2) +
  theme_light() +
  labs(title = "Lag-Corrected Alignment",
       subtitle = paste("Behavior shifted by", max_lag_s, "ms to match Neural dynamics"),
       y = "Z-scored Difference", x = "Time (s)") +
  scale_color_manual(values = c("orange","#E41A1C", "#377EB8"))

savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", height = 3,width = 6,  unit = "in")
savefig %<>%
  fill_panel(plotplot, column = 1, row = 1)
savefig

# save_multi_panel_figure(savefig, "figures/Crosscorr_Curves_RTSSVEP.pdf")

```

#### calculate cross correlation with permutation statistics

```{r calc_crosscorrelation_difference_smooths_permutation, warning = FALSE, fig.height=4, fig.width=6}

# define data

run_permutation <- FALSE  # Set to TRUE to run, FALSE to load from disk
results_filename <- "GAM_Permutation_Results_RS.RData"
n_perm <- 5000 #500 

if (run_permutation) {
  # --- 2. RUN THE LOOP ---
  cat("Running permutations... this may take a while.\n")
  
  n_perm <- 5000 #500 
  null_ccfs_RT <- matrix(NA, nrow = n_perm, ncol = 101) # 101 is default for lag.max=50
  s_beh_perm_RT  <- matrix(NA, nrow = n_perm, ncol = length(common_times)) # 101 is default for lag.max=50
  s_neu_perm_RT  <- s_beh_perm_RT
  s_beh_perm_RT_z <- s_beh_perm_RT
  s_neu_perm_RT_z <- s_beh_perm_RT
  
  # df <- beh_data
  
  
  # # troubleshooting
  # test_diff <- get_diff_smooth_vector(beh_data, is_behavioral = TRUE, time_grid = common_times)
  # test_diff <- get_diff_smooth_vector(neur_data, is_behavioral = FALSE, time_grid = common_times)
  
  for(i in 1:n_perm) {
    # Shuffle labels
    shuff_beh <- beh_data %>% 
      mutate(selection = sample(selection))
    shuff_neu <- neur_data  %>% 
      mutate(selection = sample(selection))
    
    # Extract difference smooths for the shuffled data
    s_beh <- get_smooth_vector(shuff_beh, datainput = 'RT', common_times)
    s_neu <- get_smooth_vector(shuff_neu, datainput = 'SSVEP', common_times)
    
    # Standardize (Z-score) before CCF
    s_beh_z <- as.numeric(scale(s_beh$diff))
    s_neu_z <- as.numeric(scale(s_neu$diff))
    
    # Run CCF
    res <- ccf(s_neu_z, s_beh_z, lag.max = 50, plot = FALSE)
    null_ccfs_RT[i, ] <- res$acf
    s_beh_perm_RT[i, ] <- s_beh$diff
    s_neu_perm_RT[i, ] <- s_neu$diff
    s_beh_perm_RT_z[i, ] <- s_beh_z
    s_neu_perm_RT_z[i, ] <- s_neu_z
    
    if(i %% 50 == 0) cat("Completed permutation", i, "\n")
  }
  
  # --- 3. SAVE TO DISK ---
  save(null_ccfs_RT, s_beh_perm_RT, s_neu_perm_RT, s_beh_perm_RT_z, s_neu_perm_RT_z, 
       file = results_filename)
  cat("Permutations complete. Results saved to:", results_filename, "\n")
} else {
  # --- 4. LOAD PREVIOUSLY SAVED DATA ---
  if (file.exists(results_filename)) {
    load(results_filename)
    cat("Loaded permutation results from:", results_filename, "\n")
  } else {
    stop("Switch is set to 'FALSE' but the results file was not found! Please set switch to 'TRUE' to run permutations once.")
  }
}
  # Calculate 95% Confidence Ribbon
ccf_df_RT$lower_CI <- apply(null_ccfs_RT, 2, quantile, probs = 0.025)
ccf_df_RT$upper_CI <- apply(null_ccfs_RT, 2, quantile, probs = 0.975)


```

#### plot cross correlation with statistics

```{r plot_crosscorrelation_difference_smooths_permutation, warning = FALSE, fig.height=3, fig.width=4}

# Calculate Montecarlo p-values
obs_acf_RT <- as.numeric(ccf_result_RT$acf) # Length is e.g., 101

# 2. Calculate the p-value for each lag
# We compare the absolute values (two-tailed test)
# 'abs(null_ccfs)' is [n_perm x 101]
# 'abs(obs_acf)' is [101]
# R will recycle obs_acf across the rows of null_ccfs automatically
greater_than_obs <- sweep(abs(null_ccfs_RT), 2, abs(obs_acf_RT), ">=")

# 3. Monte Carlo p-value formula: (Count + 1) / (N + 1)
# colSums gives us one count per lag
p_values <- (colSums(greater_than_obs) + 1) / (n_perm + 1)

# 4. Add to your results data frame
ccf_df_RT$p_val <- p_values


# --- 1. Prepare Permutation Data ---
# Function to convert matrix to long format
matrix_to_long <- function(mat, modality_name, time_vec) {
  as.data.frame(mat) %>%
    mutate(PermID = row_number()) %>%
    pivot_longer(cols = -PermID, names_to = "TimeBin", values_to = "Value") %>%
    mutate(
      # Convert TimeBin (e.g., "V1") to actual time values
      Time = time_vec[as.numeric(gsub("V", "", TimeBin))],
      Modality = modality_name,
      Type = "Permuted"
    )
}

perm_beh_long <- matrix_to_long(s_beh_perm_RT, "Behavior (Hit Rate)", common_times)
perm_neu_long <- matrix_to_long(s_neu_perm_RT, "Neural (SSVEP)", common_times)
perm_beh_long_z <- matrix_to_long(s_beh_perm_RT_z, "Behavior (Hit Rate)", common_times)
perm_neu_long_z <- matrix_to_long(s_neu_perm_RT_z, "Neural (SSVEP)", common_times)

# Combine them
perm_data_all <- bind_rows(perm_beh_long, perm_neu_long)
perm_data_all_z <- bind_rows(perm_beh_long_z, perm_neu_long_z)

# --- 2. Prepare Observed Data ---
obs_diff_beh_z <- as.numeric(scale(get_smooth_vector(beh_data, datainput = 'RT', common_times)$diff))
obs_diff_neu_z <- as.numeric(scale(get_smooth_vector(neur_data, datainput = 'SSVEP', common_times)$diff))
obs_diff_beh <- as.numeric(get_smooth_vector(beh_data, datainput = 'RT', common_times)$diff)
obs_diff_neu <- as.numeric(get_smooth_vector(neur_data, datainput = 'SSVEP', common_times)$diff)


obs_data_all <- data.frame(
  Time = rep(common_times, 2),
  Value_z = c(obs_diff_beh_z, obs_diff_neu_z), # Use the Z-scored versions
  Value = c(obs_diff_beh, obs_diff_neu), # Use the Z-scored versions
  Modality = rep(c("Behavior (Hit Rate)", "Neural (SSVEP)"), each = length(common_times)),
  Type = "Observed"
)

ggplot() +
  # Plot the Permuted Lines (Spaghetti)
  geom_line(data = perm_data_all, 
            aes(x = Time, y = Value, group = interaction(PermID, Modality)), 
            color = "gray80", alpha = 0.2, size = 0.5) +
  
  # Plot the Observed Line (Bold)
  geom_line(data = obs_data_all, 
            aes(x = Time, y = Value, color = Modality), 
            size = 1.2) +
  
  # Facet by Modality to see them side-by-side
  facet_wrap(~Modality, scales = "free_y") +
  
  # Styling
  theme_minimal() +
  scale_color_manual(values = c("Behavior (Hit Rate)" = "#d6604d", "Neural (SSVEP)" = "#4393c3")) +
  labs(
    title = "Observed Dynamics vs. Null Distribution",
    subtitle = "Gray lines represent 5000 permutations with shuffled condition labels",
    x = "Time (s)",
    y = "Difference Smooth in measurement units",
    color = "Modality"
  ) +
  theme(legend.position = "none") # Facet labels already identify the plots

ggplot() +
  # Plot the Permuted Lines (Spaghetti)
  geom_line(data = perm_data_all_z, 
            aes(x = Time, y = Value, group = interaction(PermID, Modality)), 
            color = "gray80", alpha = 0.2, size = 0.5) +
  
  # Plot the Observed Line (Bold)
  geom_line(data = obs_data_all, 
            aes(x = Time, y = Value_z, color = Modality), 
            size = 1.2) +
  
  # Facet by Modality to see them side-by-side
  facet_wrap(~Modality, scales = "free_y") +
  theme_minimal() +
  scale_color_manual(values = c("Behavior (Hit Rate)" = "#d6604d", "Neural (SSVEP)" = "#4393c3")) +
  labs(
    title = "Observed Dynamics vs. Null Distribution",
    subtitle = "Gray lines represent 5000 permutations with shuffled condition labels",
    x = "Time (s)",
    y = "z scored Difference Smooth",
    color = "Modality"
  ) +
  theme(legend.position = "none") # Facet labels already identify the plots



plotplot <- ggplot(ccf_df_RT, aes(x = Lag_s, y = Correlation)) +
  # The Null Distribution Ribbon (from permutations)
  geom_ribbon(aes(ymin = lower_CI, ymax = upper_CI), fill = "grey80", alpha = 0.5) +
  # Zero lines for reference
  geom_hline(yintercept = 0, color = "black", size = 0.3) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  # The observed cross-correlation
  geom_line(color = "#2166ac", size = 1) +
  # Highlight the peak
  # Highlight significant points where p < 0.05
  geom_point(data = subset(ccf_df_RT, p_val < 0.05), 
             aes(x = Lag_s, y = Correlation), 
             color = "#2166ac", size = 2) +
  geom_segment(aes(x = max_lag_s,
                 xend = max_lag_s,
                 y = 0,
                 yend = max_cor),
             color = "red",
             linewidth = 1) +
  geom_point(aes(x = max_lag_s, y = max_cor), color = "red", size = 3) +
  coord_cartesian(ylim = c(-1, 1))+
  theme_light() +
  labs(
    title = "Cross-Correlation: SSVEP vs. Behavior",
    subtitle = c(subtitle = paste("Peak correlation of", round(max_cor, 3), "at", max_lag_s, "s lag with p = ", ccf_df_RT$p_val[which.max(abs(ccf_df_RT$Correlation))],
                                  "\nGrey ribbon = 95% null distribution (permutation-based)")),
    x = "Lag (s) [Shift of Behavior relativ to SSVEP]",
    y = "Cross-Correlation (r)"
  )


savefig <- multi_panel_figure(columns = 1, rows = 1, panel_label_type = "none", height = 3,width = 4,  unit = "in")
savefig %<>%
  fill_panel(plotplot, column = 1, row = 1)
savefig

# save_multi_panel_figure(savefig, "figures/Crosscorr_Histogram_RTSSVEP.pdf")
```
